---
title: "Supplement B - Simulation Atlas"
subtitle: "Extended sensitivity analyses for The Link Function Problem in Psychological Interaction Testing"
author:
  - Margherita Calderan
  - Filippo Gambarota
  - Laura Sità
  - Tommaso Feraco
  - Enrico Toffalini
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    df-print: paged
  pdf:
    toc: true
    number-sections: true
    include-in-header:
      text: |
        % Table typography: wrapped lines inside a cell are set slightly tighter
        % than normal and each row gets a little extra vertical air, so a line
        % break inside a cell is not mistaken for the start of a new table row.
        \usepackage{etoolbox}
        \usepackage{array}
        \AtBeginEnvironment{longtable}{%
          \linespread{0.95}\selectfont
          \setlength{\extrarowheight}{4pt}}
        \AtBeginEnvironment{tabular}{%
          \linespread{0.95}\selectfont
          \setlength{\extrarowheight}{4pt}}
execute:
  echo: false
  warning: false
  message: false
editor: source
---

```{r}
#| label: setup
#| include: false

if (!requireNamespace("ggplot2", quietly = TRUE)) {
  stop("Package 'ggplot2' is required to render the Simulation Atlas.")
}
if (!requireNamespace("knitr", quietly = TRUE)) {
  stop("Package 'knitr' is required to render the Simulation Atlas.")
}

library(ggplot2)

# Locate simulation-atlas/data whether this file is rendered from the
# repository root or from simulation-atlas/ itself.
input_file <- tryCatch(knitr::current_input(), error = function(e) "")
input_dir <- if (nzchar(input_file)) dirname(normalizePath(input_file, mustWork = FALSE)) else "."

candidate_roots <- unique(c(
  ".",
  "simulation-atlas",
  input_dir,
  file.path(input_dir, "simulation-atlas")
))

has_atlas_data <- vapply(
  candidate_roots,
  function(root) file.exists(file.path(root, "data", "atlas-summary.csv")),
  logical(1)
)

if (!any(has_atlas_data)) {
  stop(
    "Could not find simulation-atlas/data/atlas-summary.csv. ",
    "Render this document from the repository root or place it inside simulation-atlas/."
  )
}

atlas_root <- candidate_roots[which(has_atlas_data)[1]]

data_path <- function(name) file.path(atlas_root, "data", name)

core <- utils::read.csv(
  data_path("atlas-summary.csv"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

diagnostic <- utils::read.csv(
  data_path("diagnostic-atlas-summary.csv"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

grid <- utils::read.csv(
  data_path("scenario-grid.csv"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

# Basic integrity checks. The defining condition of the atlas is that the
# generating interaction equals zero in every scenario.
stopifnot(
  nrow(core) > 0,
  nrow(grid) > 0,
  all(is.finite(grid$generating_interaction)),
  all(grid$generating_interaction == 0),
  !anyDuplicated(grid$scenario_id)
)

if (!all(core$run_type == "full")) {
  warning("atlas-summary.csv is not entirely from a full run.")
}

pct <- function(x, digits = 1) {
  ifelse(is.finite(x), paste0(formatC(100 * x, format = "f", digits = digits), "%"), "--")
}

num <- function(x, digits = 3) {
  ifelse(is.finite(x), formatC(x, format = "f", digits = digits), "--")
}

ci_text <- function(low, high, digits = 1) {
  paste0("[", pct(low, digits), ", ", pct(high, digits), "]")
}

# Slice identifiers are code-level values that use underscores. Underscores are
# unbreakable in the PDF tables, so display them with spaces instead.
slice_display <- function(x) gsub("_", " ", x, fixed = TRUE)

has_slice <- function(x, slice) {
  vapply(
    strsplit(ifelse(is.na(x), "", x), ";", fixed = TRUE),
    function(z) slice %in% z,
    logical(1)
  )
}

family_title <- c(
  forced_choice = "Simulation 1: Forced-choice accuracy and the non-zero chance floor",
  within_family = "Simulation 2: Within-family link choice, logit versus probit",
  sum_scores = "Simulation 3: Sum scores"
)

family_short <- c(
  forced_choice = "Forced-choice accuracy",
  within_family = "Logit vs probit",
  sum_scores = "Sum scores"
)

model_order <- list(
  forced_choice = c(
    "Chance-corrected binomial logit",
    "Standard binomial logit",
    "Standard binomial probit",
    "Gaussian identity"
  ),
  within_family = c("Binomial logit", "Binomial probit"),
  sum_scores = c(
    "Latent generating scale",
    "Gaussian-probit bounded score",
    "Observed sum-score identity"
  )
)

slice_specs <- list(
  forced_choice = data.frame(
    slice = c("sample_size", "scale_location", "trial_count", "chance_level"),
    label = c("Sample size", "Intercept", "Trials per participant", "Chance level"),
    variable = c("N", "beta_intercept", "k_trials", "chance"),
    xlab = c("N", "Generating intercept", "Trials per participant", "Chance level"),
    stringsAsFactors = FALSE
  ),
  within_family = data.frame(
    slice = c("sample_size", "trial_count", "latent_location", "icc"),
    label = c("Sample size", "Trials per cell", "Intercept", "Latent ICC"),
    variable = c("n_subjects", "k_trials", "reference_beta_intercept", "target_icc"),
    xlab = c("Participants", "Trials per participant-condition cell", "Reference intercept", "Latent ICC"),
    stringsAsFactors = FALSE
  ),
  sum_scores = data.frame(
    slice = c("sample_size", "scale_location", "item_count", "latent_dispersion"),
    label = c("Sample size", "Scale location", "Number of items", "Latent dispersion"),
    variable = c("N", "threshold_shift", "J", "theta_sd"),
    xlab = c("N", "Item-threshold shift", "Number of items", "Latent SD"),
    stringsAsFactors = FALSE
  )
)

family_rows <- function(family) {
  out <- core[core$family == family, , drop = FALSE]
  out$model_label <- factor(out$model_label, levels = model_order[[family]])
  out
}

anchor_rows <- function(family) {
  out <- family_rows(family)
  out[out$paper_anchor %in% TRUE, , drop = FALSE]
}

slice_rows <- function(family) {
  specs <- slice_specs[[family]]
  dat <- family_rows(family)
  pieces <- lapply(seq_len(nrow(specs)), function(i) {
    spec <- specs[i, ]
    keep <- has_slice(dat$slice_membership, spec$slice)
    out <- dat[keep, , drop = FALSE]
    out$slice_view <- spec$slice
    out$slice_label <- spec$label
    out$x_value <- suppressWarnings(as.numeric(out[[spec$variable]]))
    out$x_label <- spec$xlab
    out
  })
  do.call(rbind, pieces)
}

surface_rows <- function(family) {
  dat <- family_rows(family)
  dat[has_slice(dat$slice_membership, "main_effect_surface"), , drop = FALSE]
}

anchor_table <- function(family) {
  dat <- anchor_rows(family)
  out <- data.frame(
    Scenario = dat$scenario_label,
    `Generating link/scale` = dat$generating_link,
    `Fitted model/scale` = as.character(dat$model_label),
    `Rate type` = dat$rate_type,
    `Rejection rate` = pct(dat$false_positive_rate),
    `95% CI` = ci_text(dat$false_positive_ci_low, dat$false_positive_ci_high),
    `Deterministic pseudo-interaction` = num(dat$deterministic_pseudo_interaction),
    check.names = FALSE
  )
  rownames(out) <- NULL
  out
}

plot_anchor_rates <- function(family) {
  dat <- anchor_rows(family)
  dat$model_label <- factor(dat$model_label, levels = model_order[[family]])

  ggplot(
    dat,
    aes(
      x = model_label,
      y = false_positive_rate,
      ymin = false_positive_ci_low,
      ymax = false_positive_ci_high,
      group = scenario_label,
      shape = scenario_label
    )
  ) +
    geom_hline(yintercept = 0.05, linetype = "dashed", linewidth = 0.4) +
    geom_pointrange(position = position_dodge(width = 0.55)) +
    scale_y_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.2),
      labels = function(x) paste0(round(100 * x), "%")
    ) +
    labs(
      x = NULL,
      y = "Product-term rejection rate",
      shape = "Scenario"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 25, hjust = 1),
      legend.position = "top"
    )
}

plot_slice_rates <- function(family) {
  dat <- slice_rows(family)
  dat$model_label <- factor(dat$model_label, levels = model_order[[family]])
  dat$slice_label <- factor(dat$slice_label, levels = slice_specs[[family]]$label)

  p <- ggplot(
    dat,
    aes(
      x = x_value,
      y = false_positive_rate,
      group = model_label,
      linetype = model_label,
      shape = model_label
    )
  ) +
    geom_hline(yintercept = 0.05, linetype = "dashed", linewidth = 0.35) +
    geom_line() +
    geom_point(size = 2) +
    scale_y_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.2),
      labels = function(x) paste0(round(100 * x), "%")
    ) +
    labs(
      x = NULL,
      y = "Product-term rejection rate",
      linetype = "Fitted model",
      shape = "Fitted model"
    ) +
    theme_minimal(base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "top"
    )

  if (family == "within_family") {
    p + facet_grid(generating_link ~ slice_label, scales = "free_x")
  } else {
    p + facet_wrap(~ slice_label, scales = "free_x", ncol = 2)
  }
}

slice_range_table <- function(family) {
  dat <- slice_rows(family)
  key <- interaction(dat$slice_label, dat$generating_link, dat$model_label, drop = TRUE)
  split_dat <- split(dat, key)

  rows <- lapply(split_dat, function(x) {
    data.frame(
      Slice = x$slice_label[1],
      `Generating link/scale` = x$generating_link[1],
      `Fitted model/scale` = as.character(x$model_label[1]),
      `Minimum rejection rate` = pct(min(x$false_positive_rate, na.rm = TRUE)),
      `Maximum rejection rate` = pct(max(x$false_positive_rate, na.rm = TRUE)),
      check.names = FALSE
    )
  })

  out <- do.call(rbind, rows)
  rownames(out) <- NULL
  out
}

coefficient_values_text <- function(family, variable, generating_link = NULL, digits = 3) {
  dat <- surface_rows(family)
  if (!is.null(generating_link)) {
    dat <- dat[dat$generating_link == generating_link, , drop = FALSE]
  }
  values <- sort(unique(dat[[variable]][is.finite(dat[[variable]])]))
  paste(formatC(values, format = "f", digits = digits), collapse = ", ")
}

main_effect_coefficient_table <- function() {
  rows <- list(
    data.frame(
      `Manuscript case` = "Simulation 1: Forced-choice accuracy",
      `Generating link/scale` = "Chance-corrected logit",
      `First main effect` = "Age coefficient",
      `Coefficient values` = coefficient_values_text("forced_choice", "beta_age", digits = 2),
      `Second main effect` = "Group coefficient",
      `Coefficient values 2` = coefficient_values_text("forced_choice", "beta_group", digits = 2),
      check.names = FALSE
    ),
    data.frame(
      `Manuscript case` = "Simulation 2: Logit vs probit",
      `Generating link/scale` = "Logit",
      `First main effect` = "Group coefficient",
      `Coefficient values` = coefficient_values_text("within_family", "beta_group", "logit", digits = 3),
      `Second main effect` = "Condition coefficient",
      `Coefficient values 2` = coefficient_values_text("within_family", "beta_condition", "logit", digits = 3),
      check.names = FALSE
    ),
    data.frame(
      `Manuscript case` = "Simulation 2: Logit vs probit",
      `Generating link/scale` = "Probit",
      `First main effect` = "Group coefficient",
      `Coefficient values` = coefficient_values_text("within_family", "beta_group", "probit", digits = 2),
      `Second main effect` = "Condition coefficient",
      `Coefficient values 2` = coefficient_values_text("within_family", "beta_condition", "probit", digits = 2),
      check.names = FALSE
    ),
    data.frame(
      `Manuscript case` = "Simulation 3: Sum scores",
      `Generating link/scale` = "Latent generating scale",
      `First main effect` = "Predictor coefficient",
      `Coefficient values` = coefficient_values_text("sum_scores", "beta_x", digits = 3),
      `Second main effect` = "Group coefficient",
      `Coefficient values 2` = coefficient_values_text("sum_scores", "beta_group", digits = 2),
      check.names = FALSE
    )
  )
  out <- do.call(rbind, rows)
  names(out)[4] <- "First coefficient values"
  names(out)[6] <- "Second coefficient values"
  rownames(out) <- NULL
  out
}

plot_surface <- function(family) {
  dat <- surface_rows(family)
  dat$model_label <- factor(dat$model_label, levels = model_order[[family]])

  if (family == "forced_choice") {
    dat$surface_x <- factor(dat$beta_age, levels = sort(unique(dat$beta_age)))
    dat$surface_y <- factor(dat$beta_group, levels = sort(unique(dat$beta_group)))
    xlab <- "Age coefficient on generating scale"
    ylab <- "Group coefficient on generating scale"
  } else if (family == "within_family") {
    dat$surface_x <- factor(dat$beta_group, levels = sort(unique(dat$beta_group)))
    dat$surface_y <- factor(dat$beta_condition, levels = sort(unique(dat$beta_condition)))
    xlab <- "Group coefficient on generating link scale"
    ylab <- "Condition coefficient on generating link scale"
  } else {
    dat$surface_x <- factor(dat$beta_x, levels = sort(unique(dat$beta_x)))
    dat$surface_y <- factor(dat$beta_group, levels = sort(unique(dat$beta_group)))
    xlab <- "Predictor coefficient on latent generating scale"
    ylab <- "Group coefficient on latent generating scale"
  }

  p <- ggplot(
    dat,
    aes(
      x = surface_x,
      y = surface_y,
      fill = false_positive_rate
    )
  ) +
    geom_tile() +
    geom_text(aes(label = paste0(round(100 * false_positive_rate), "%")), size = 3, colour = "white") +
    scale_fill_gradient(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.25),
      labels = function(x) paste0(round(100 * x), "%")
    ) +
    labs(
      x = xlab,
      y = ylab,
      fill = "Rejection rate"
    ) +
    theme_minimal(base_size = 10) +
    theme(panel.grid = element_blank(), legend.position = "right")

  if (family == "within_family") {
    p + facet_wrap(vars(generating_link, model_label), ncol = 2, scales = "free")
  } else {
    p + facet_wrap(~ model_label)
  }
}

rate_range <- function(family, model = NULL, rate_type = NULL) {
  dat <- family_rows(family)
  if (!is.null(model)) dat <- dat[as.character(dat$model_label) == model, , drop = FALSE]
  if (!is.null(rate_type)) dat <- dat[dat$rate_type == rate_type, , drop = FALSE]
  range(dat$false_positive_rate, na.rm = TRUE)
}

rate_range_text <- function(...) {
  r <- rate_range(...)
  paste0(pct(r[1]), " to ", pct(r[2]))
}

# Compact scenario-level appendix. Keep only columns needed to reconstruct the
# scientific result without reproducing the wide machine-readable CSV.
compact_core_table <- function(family) {
  dat <- family_rows(family)
  out <- data.frame(
    ID = dat$scenario_id,
    Slice = slice_display(dat$slice),
    Scenario = dat$scenario_label,
    `Generating link/scale` = dat$generating_link,
    `Fitted model/scale` = as.character(dat$model_label),
    `Rejection rate` = pct(dat$false_positive_rate),
    `Deterministic pseudo-interaction` = num(dat$deterministic_pseudo_interaction),
    check.names = FALSE
  )
  rownames(out) <- NULL
  out
}

report_kable <- function(x, align = NULL, long = FALSE, ...) {
  if (long && knitr::is_latex_output()) {
    knitr::kable(x, align = align, longtable = TRUE, ...)
  } else {
    knitr::kable(x, align = align, ...)
  }
}

# Diagnostic helpers -------------------------------------------------------

diagnostic_family_title <- c(
  forced_choice = "Forced-choice chance-floor case",
  within_family = "Probit-generated, logit-fitted case"
)

diagnostic_expand_slices <- function(dat) {
  pieces <- list()
  counter <- 1L

  for (slice in c("sample_size", "severity")) {
    keep <- has_slice(dat$diagnostic_slice_membership, slice)
    out <- dat[keep, , drop = FALSE]
    if (!nrow(out)) next

    out$slice_view <- slice
    out$slice_label <- if (slice == "sample_size") "Sample size" else "Intercept"
    out$x_value <- NA_real_

    if (slice == "sample_size") {
      out$x_value[out$family == "forced_choice"] <- out$N[out$family == "forced_choice"]
      out$x_value[out$family == "within_family"] <- out$n_subjects[out$family == "within_family"]
    } else {
      out$x_value[out$family == "forced_choice"] <- out$beta_intercept[out$family == "forced_choice"]
      out$x_value[out$family == "within_family"] <- out$reference_beta_intercept[out$family == "within_family"]
    }

    pieces[[counter]] <- out
    counter <- counter + 1L
  }

  do.call(rbind, pieces)
}

make_diagnostic_plot_data <- function() {
  base <- diagnostic[diagnostic$family %in% c("forced_choice", "within_family"), , drop = FALSE]
  dat <- diagnostic_expand_slices(base)
  dat$family_label <- unname(diagnostic_family_title[dat$family])

  # One pseudo-interaction row per scenario within each slice. The manuscript
  # anchor belongs to both sensitivity slices and is therefore shown in both.
  pseudo_key <- interaction(dat$scenario_id, dat$slice_view, drop = TRUE)
  pseudo <- dat[!duplicated(pseudo_key), c(
    "scenario_id", "family", "family_label", "slice_label", "x_value",
    "pseudo_interaction_detection_rate", "pseudo_interaction_ci_low",
    "pseudo_interaction_ci_high"
  )]
  names(pseudo)[names(pseudo) == "pseudo_interaction_detection_rate"] <- "rate"
  names(pseudo)[names(pseudo) == "pseudo_interaction_ci_low"] <- "low"
  names(pseudo)[names(pseudo) == "pseudo_interaction_ci_high"] <- "high"
  pseudo$quantity <- "Pseudo-interaction detection"

  # Retain diagnostics that were used for the relevant scenario.
  checks <- dat[dat$applicable %in% TRUE & dat$computed %in% TRUE, , drop = FALSE]
  checks <- checks[, c(
    "scenario_id", "family", "family_label", "slice_label", "x_value",
    "diagnostic", "detection_rate", "detection_ci_low", "detection_ci_high"
  )]
  names(checks)[names(checks) == "detection_rate"] <- "rate"
  names(checks)[names(checks) == "detection_ci_low"] <- "low"
  names(checks)[names(checks) == "detection_ci_high"] <- "high"
  names(checks)[names(checks) == "diagnostic"] <- "quantity"

  common <- c("scenario_id", "family", "family_label", "slice_label", "x_value", "quantity", "rate", "low", "high")
  rbind(pseudo[, common], checks[, common])
}

plot_diagnostic_atlas <- function() {
  dat <- make_diagnostic_plot_data()

  ggplot(
    dat,
    aes(
      x = x_value,
      y = rate,
      group = quantity,
      linetype = quantity,
      shape = quantity
    )
  ) +
    geom_line() +
    geom_point(size = 2) +
    scale_y_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.2),
      labels = function(x) paste0(round(100 * x), "%")
    ) +
    facet_grid(family_label ~ slice_label, scales = "free_x") +
    labs(
      x = NULL,
      y = "Detection / selection rate",
      linetype = NULL,
      shape = NULL
    ) +
    theme_minimal(base_size = 10) +
    theme(panel.grid.minor = element_blank(), legend.position = "top")
}

aic_table <- function() {
  base <- diagnostic[
    diagnostic$family %in% c("forced_choice", "within_family") &
      diagnostic$diagnostic == "AIC favors generating link",
    ,
    drop = FALSE
  ]
  dat <- diagnostic_expand_slices(base)

setting_value <- ifelse(
  dat$slice_label == "Sample size",
  formatC(dat$x_value, format = "f", digits = 0),
  formatC(dat$x_value, format = "f", digits = 2)
)

out <- data.frame(
  Case = unname(diagnostic_family_title[dat$family]),
  Setting = paste0(dat$slice_label, " = ", setting_value),
  `Pseudo-interaction detection` = pct(dat$pseudo_interaction_detection_rate),
  `AIC favors generating link` = pct(dat$detection_rate),
  `Median delta AIC` = num(dat$aic_delta_median, 2),
  `Delta AIC IQR` = paste0("[", num(dat$aic_delta_q25, 2), ", ", num(dat$aic_delta_q75, 2), "]"),
  `Absolute delta AIC < 2` = pct(dat$aic_delta_within_two_rate),
  check.names = FALSE
)
  
  rownames(out) <- NULL
  out
}

diagnostic_used_table <- function() {
  dat <- diagnostic[
    diagnostic$family %in% c("forced_choice", "within_family") &
      diagnostic$applicable %in% TRUE & diagnostic$computed %in% TRUE,
    ,
    drop = FALSE
  ]
  parts <- split(dat, dat$family)
  out <- do.call(rbind, lapply(parts, function(x) {
    data.frame(
      `Diagnostic case` = unname(diagnostic_family_title[x$family[1]]),
      `Diagnostics used` = paste(unique(x$diagnostic), collapse = "; "),
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  }))
  rownames(out) <- NULL
  out
}

scenario_counts <- table(grid$family)
```

# Purpose and scope

Supplement B extends the three manuscript simulations by varying selected parameters while keeping the true interaction equal to zero on the generating scale. It uses one-dimensional sensitivity slices and small main-effect surfaces rather than a full Cartesian design.

Reference scenarios retain the generating settings of the corresponding manuscript simulations. In every scenario, a rejection under the matched generating model estimates the nominal generating-scale rejection rate; under another scale, it is a **pseudo-interaction detection rate**. All scenarios used 3,000 Monte Carlo replications.

## How to read the figures

All rejection-rate figures use the same interpretation. The horizontal dashed line at .05 marks nominal alpha. Rates close to .05 under the generating scale provide a calibration check. Elevated rates under an alternative fitted scale show how often the scale difference becomes a statistically significant product term at the stated sample size and design. Error bars are shown for the reference scenarios. The sensitivity plots omit error bars to keep the larger set of scenarios readable; Wilson 95% intervals are available in the numerical tables and in `data/atlas-summary.csv`.

The deterministic pseudo-interaction reported in the tables is calculated from expected values before sampling. It describes the interaction contrast implied by evaluating the generating pattern on the fitted scale. The Monte Carlo rejection rate adds sampling variability, standard errors, and sample size to that deterministic discrepancy.

# Atlas design

```{r}
#| label: tbl-atlas-design
#| tbl-cap: "Structure of the simulation atlas. Each family contains reference scenarios plus prespecified sensitivity slices."

design_table <- data.frame(
  `Manuscript case` = c(
    "Simulation 1: Forced-choice accuracy",
    "Simulation 2: Logit vs probit",
    "Simulation 3: Sum scores"
  ),
  `Generating scenarios` = c(
    scenario_counts[["forced_choice"]],
    scenario_counts[["within_family"]],
    scenario_counts[["sum_scores"]]
  ),
  `Sensitivity slices` = c(
    "Main effects; sample size; intercept; trials; chance level",
    "Main effects; sample size; trials; intercept; ICC",
    "Main effects; sample size; scale location; item count; latent dispersion"
  ),
  check.names = FALSE
)

knitr::kable(design_table, align = c("l", "r", "l"))
```

The one-dimensional slices change one named parameter while holding the remaining settings at the central reference scenario. The main-effect surfaces vary the two additive main-effect coefficients jointly. The actual coefficients are reported on the figure axes and in the table below.

```{r}
#| label: tbl-atlas-main-effect-coefficients
#| tbl-cap: "Generating-scale coefficients used in the main-effect surfaces. Coefficients are reported on the scale that generated each scenario."

knitr::kable(main_effect_coefficient_table(), align = c("l", "l", "l", "l", "l", "l"))
```

For Simulation 2, logit and probit coefficients are on different generating-link scales. As in the manuscript simulation, the logit coefficients are scaled by 1.65 relative to the probit reference coefficients so that the two generating scenarios have similar cell probabilities. The figures therefore report the actual coefficient values for each generating link instead of a common relative index.


# Simulation 1: Forced-choice accuracy and the non-zero chance floor

Forced-choice accuracy is generated from a chance-corrected binomial logit model with a .50 floor and no age-by-group product term on that scale. The fitted models are the matched chance-corrected model, standard binomial logit and probit models, and a Gaussian identity model; the figures and Table 1 show the sensitivity design.

## Reference scenarios

```{r}
#| label: fig-atlas-fc-anchor
#| fig-cap: "Forced-choice reference scenarios. The product term is zero on the chance-corrected logit generating scale. The dashed line marks alpha = .05."
#| fig-width: 8
#| fig-height: 4.8

plot_anchor_rates("forced_choice")
```

```{r}
#| label: tbl-atlas-fc-anchor
#| tbl-cap: "Forced-choice reference scenarios. The deterministic pseudo-interaction is evaluated on each fitted scale before sampling."

knitr::kable(anchor_table("forced_choice"), align = c("l", "l", "l", "l", "r", "r", "r"))
```

Across all forced-choice atlas scenarios, the matched chance-corrected model remains close to nominal alpha, with rejection rates from `r rate_range_text("forced_choice", model = "Chance-corrected binomial logit")`. The standard binomial logit ranges from `r rate_range_text("forced_choice", model = "Standard binomial logit")`, and the standard binomial probit from `r rate_range_text("forced_choice", model = "Standard binomial probit")`. These ranges are sensitivity results, not prevalence estimates. They show that the inferential consequence of placing the lower asymptote at 0 instead of the task-implied chance level can be small in some regions and very large in others.

## One-dimensional sensitivity slices

```{r}
#| label: fig-atlas-fc-slices
#| fig-cap: "Forced-choice sensitivity slices. One parameter varies at a time around the central reference scenario. The dashed line marks alpha = .05."
#| fig-width: 9
#| fig-height: 7

plot_slice_rates("forced_choice")
```

```{r}
#| label: tbl-atlas-fc-slice-ranges
#| tbl-cap: "Range of rejection rates across each forced-choice sensitivity slice."

knitr::kable(slice_range_table("forced_choice"), align = c("l", "l", "l", "r", "r"))
```

The slices separate distinct reasons why a deterministic scale discrepancy may or may not become a statistical interaction. The intercept and the chance level change the curvature relevant to the fitted link. Sample size and number of trials change how precisely that discrepancy is estimated. The figure therefore distinguishes the existence of a pseudo-interaction on the fitted scale from the probability that a product-term test detects it.

## Main-effect surface

```{r}
#| label: fig-atlas-fc-surface
#| fig-cap: "Forced-choice main-effect surface. Axes report the age and group coefficients on the chance-corrected logit generating scale. Cell labels are product-term rejection rates."
#| fig-width: 9
#| fig-height: 5.8

plot_surface("forced_choice")
```

The surface directly targets the mechanism emphasized in the manuscript. Two additive main effects can jointly move observations through regions where equal differences on the generating scale map to unequal differences on the fitted scale. Larger main effects can therefore increase the deterministic pseudo-interaction even though the generating age-by-group term remains zero.

# Simulation 2: Within-family link choice, logit versus probit

Repeated-trial random-intercept GLMMs cross logit and probit generating and fitted links. Both use the binomial family and the same 0-to-1 bounds, but impose additivity on different scales; the figures and Table 1 show the sensitivity design.

## Reference scenarios

```{r}
#| label: fig-atlas-wf-anchor
#| fig-cap: "Logit-probit reference scenarios. The product term is zero on the generating link scale. The dashed line marks alpha = .05."
#| fig-width: 8
#| fig-height: 4.8

plot_anchor_rates("within_family")
```

```{r}
#| label: tbl-atlas-wf-anchor
#| tbl-cap: "Within-family reference scenarios for logit and probit generating links."

knitr::kable(anchor_table("within_family"), align = c("l", "l", "l", "l", "r", "r", "r"))
```

Across the atlas, matched logit/probit fits remain near the nominal rejection rate. The wrong-link rows range from `r pct(min(family_rows("within_family")$false_positive_rate[family_rows("within_family")$rate_type == "Pseudo-interaction detection rate"], na.rm = TRUE))` to `r pct(max(family_rows("within_family")$false_positive_rate[family_rows("within_family")$rate_type == "Pseudo-interaction detection rate"], na.rm = TRUE))`. The range is wider than the reference scenarios alone suggest because the atlas deliberately varies information and the region of the latent scale. The result remains conditional on these declared slices and should not be read as a general frequency for logit-probit disagreement.

## One-dimensional sensitivity slices

```{r}
#| label: fig-atlas-wf-slices
#| fig-cap: "Logit-probit sensitivity slices. Rows identify the generating link. Within each panel the fitted logit and probit models are compared as one parameter varies around the reference scenario."
#| fig-width: 10
#| fig-height: 7

plot_slice_rates("within_family")
```

```{r}
#| label: tbl-atlas-wf-slice-ranges
#| tbl-cap: "Range of rejection rates across each within-family sensitivity slice."

knitr::kable(slice_range_table("within_family"), align = c("l", "l", "l", "r", "r"))
```

Sample size and trial count mainly change how detectable the small systematic discrepancy between the links becomes. The intercept and the main-effect magnitudes change the discrepancy itself by moving cell probabilities into regions where logit and probit curvature differs more. ICC changes the information structure of the repeated observations while leaving the fixed generating interaction at zero.

## Main-effect surface

```{r}
#| label: fig-atlas-wf-surface
#| fig-cap: "Within-family main-effect surfaces. Axes report the group and condition coefficients on each generating-link scale. Panels identify the generating and fitted links. Cell labels are product-term rejection rates."
#| fig-width: 8.5
#| fig-height: 6.5

plot_surface("within_family")
```

The matched-link panels provide the calibration reference. The off-diagonal panels show when the small difference between two standard binomial links becomes consequential for an interaction claim. This is the narrow point of the within-family example: logit and probit can remain very similar for prediction while defining sufficiently different additivity baselines to change a product-term conclusion under high information or strong curvature.

# Simulation 3: Sum scores

Sum-score data are generated on a latent continuous scale with no x-by-group interaction, converted to ordinal item responses, and summed. Analyses compare the latent generating scale, observed sum-score identity model, and Gaussian-probit bounded-score sensitivity model; the figures and Table 1 show the sensitivity design.

## Reference scenarios

```{r}
#| label: fig-atlas-ss-anchor
#| fig-cap: "Sum-score reference scenarios. The product term is zero on the latent generating scale. The dashed line marks alpha = .05."
#| fig-width: 8
#| fig-height: 4.8

plot_anchor_rates("sum_scores")
```

```{r}
#| label: tbl-atlas-ss-anchor
#| tbl-cap: "Sum-score reference scenarios. The latent generating scale is available only in simulation."

knitr::kable(anchor_table("sum_scores"), align = c("l", "l", "l", "l", "r", "r", "r"))
```

Across the atlas, the latent generating scale rejects from `r rate_range_text("sum_scores", model = "Latent generating scale")`. The observed sum-score identity model ranges from `r rate_range_text("sum_scores", model = "Observed sum-score identity")`, while the Gaussian-probit bounded-score sensitivity model ranges from `r rate_range_text("sum_scores", model = "Gaussian-probit bounded score")`. The latter remains a model of the manifest total. It does not recover the latent construct and should be read as a sensitivity analysis for the scale on which additivity is imposed.

## One-dimensional sensitivity slices

```{r}
#| label: fig-atlas-ss-slices
#| fig-cap: "Sum-score sensitivity slices. One parameter varies at a time around the central reference scenario. The dashed line marks alpha = .05."
#| fig-width: 9
#| fig-height: 7

plot_slice_rates("sum_scores")
```

```{r}
#| label: tbl-atlas-ss-slice-ranges
#| tbl-cap: "Range of rejection rates across each sum-score sensitivity slice."

knitr::kable(slice_range_table("sum_scores"), align = c("l", "l", "l", "r", "r"))
```

The scale-location slice changes how strongly the bounded sum score compresses the latent dimension. Item count and latent dispersion change the mapping between the latent variable and the manifest total. Sample size changes the probability that any resulting fitted-scale interaction is detected. These are distinct components of the sum-score problem and should not be reduced to a generic floor-or-ceiling warning.

## Main-effect surface

```{r}
#| label: fig-atlas-ss-surface
#| fig-cap: "Sum-score main-effect surface. Axes report the predictor and group coefficients on the latent generating scale. Cell labels are product-term rejection rates."
#| fig-width: 8.5
#| fig-height: 5.8

plot_surface("sum_scores")
```

The surface shows how the mapping from latent values to a bounded total can turn two additive latent effects into a non-additive pattern on the manifest score. The latent benchmark remains the calibration reference because its product term is zero by construction.

# Diagnostic atlas

The diagnostic analyses cover the forced-choice chance-floor case and the probit-generated, logit-fitted repeated-trial case. They examine how pseudo-interaction risk and diagnostic detection change with sample size and intercept.

Diagnostics are applied to the **misspecified interaction model**, matching the manuscript's post-estimation question: once the fitted model has produced the product-term finding, does that same model show enough evidence of misspecification to warn the analyst? AIC compares prespecified candidate mean models with the same structural interaction formula. Positive delta AIC, defined as AIC of the misspecified model minus AIC of the generating model, favors the generating link.

```{r}
#| label: tbl-atlas-diagnostic-status
#| tbl-cap: "Diagnostics used for each diagnostic case."

knitr::kable(diagnostic_used_table(), align = c("l", "l"))
```

Checks not listed were not applicable to the design. In particular, the design-cell residual check is used only for the categorical repeated-trial case, whereas residual quantile checks over a focal predictor require a continuous predictor.

## Detection across sensitivity slices

```{r}
#| label: fig-atlas-diagnostics
#| fig-cap: "Diagnostic atlas. Pseudo-interaction detection is shown together with each diagnostic or candidate-model comparison. Sample-size and scale-location slices are shown separately."
#| fig-width: 10
#| fig-height: 7

plot_diagnostic_atlas()
```

The rates in this figure answer different questions. Pseudo-interaction detection is the probability that the misspecified fitted scale yields a significant product term. Pregibon-style detection is the probability that the added-term link check flags a departure. AIC reports how often the generating link has the lower AIC within the specified candidate pair. The curves therefore need not be complements of one another. Diagnostic misalignment occurs when the interaction conclusion changes more readily than the diagnostic workflow signals a problem.

## Strength of the AIC comparison

```{r}
#| label: tbl-atlas-aic
#| tbl-cap: "AIC comparison across the diagnostic atlas. Positive delta AIC favors the generating link. The within-two column reports the proportion of finite comparisons with absolute delta AIC below 2."

knitr::kable(aic_table(), align = c("l", "l", "r", "r", "r", "r", "r"))
```

The AIC win rate is reported together with the magnitude of the difference because these describe different situations. A generating link that wins frequently by a very small margin is weaker evidence than one that wins by a large margin. As in the manuscript, these comparisons rank a limited candidate set. They do not identify the substantively correct scale by themselves.

# Scope and interpretation

The sensitivity results show whether the reference scenarios reflect broader patterns or depend on a narrow configuration. They also distinguish parameters that change the **deterministic scale discrepancy**, such as the intercept and main-effect magnitude, from parameters that mainly change **detectability**, such as sample size and trial count. The diagnostic analyses place pseudo-interaction detection and model-checking performance on the same sensitivity slices.

The atlas does not estimate how often link misspecification occurs in psychological research. It does not cover every plausible outcome family, link, random-effects structure, measurement model, or predictor distribution. It also does not turn model diagnostics into a decision rule for the scientifically correct scale. Its role is narrower: to show how the mechanisms developed in the paper change across a declared set of nearby scenarios.

# Numerical appendix

The tables below provide scenario-level results for the core sensitivity analyses. All scenarios used 3,000 Monte Carlo replications.

## Forced-choice accuracy

```{r}
#| label: tbl-atlas-fc-full
#| tbl-cap: "Scenario-level results for the forced-choice atlas."

report_kable(compact_core_table("forced_choice"), align = c("l", "l", "l", "l", "l", "r", "r"), long = TRUE)
```

## Logit versus probit

```{r}
#| label: tbl-atlas-wf-full
#| tbl-cap: "Scenario-level results for the within-family logit-probit atlas."

report_kable(compact_core_table("within_family"), align = c("l", "l", "l", "l", "l", "r", "r"), long = TRUE)
```

## Sum scores

```{r}
#| label: tbl-atlas-ss-full
#| tbl-cap: "Scenario-level results for the sum-score atlas."

report_kable(compact_core_table("sum_scores"), align = c("l", "l", "l", "l", "l", "r", "r"), long = TRUE)
```
