---
title: "The Link Function Problem in Psychological Interaction Testing"
shorttitle: "Link Functions and Interactions"
author:
  - name: Margherita Calderan
    orcid: 0009-0005-5668-5162
    email: margherita.calderan@unipd.it
    affiliations:
      - name: University of Padova
        department: Department of General Psychology
        address: Via Venezia 8
        city: Padova
        country: Italy
        postal-code: 35131
        id: dpg
  - name: Filippo Gambarota
    orcid: 0000-0002-6666-1747
    email: filippo.gambarota@unipd.it
    affiliations:
      - name: University of Padova
        department: Department of Developmental Psychology and Socialization
        address: Via Venezia 8
        city: Padova
        country: Italy
        postal-code: 35131
        id: dpss
  - name: Laura Sità
    orcid: 0009-0000-6017-9979
    email: laura.sita@unipd.it
    affiliations:
      - ref: dpg
  - name: Tommaso Feraco
    orcid: 0000-0002-8920-5330
    email: tommaso.feraco@unipd.it
    affiliations:
      - ref: dpg
  - name: Enrico Toffalini
    corresponding: true
    orcid: 0000-0002-1404-5133
    email: enrico.toffalini@unipd.it
    affiliations:
      - ref: dpg
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "Equal differences on one scale become different differences on another scale. This is crucial for interaction testing because an interaction is a difference between differences, and in psychological research the observed response and the construct of interest are often expressed on different scales. Therefore, the same data can show an interaction on one scale and not on another. Generalized Linear Models make this problem explicit by separating two decisions, the family, which describes the nature of the response variable, and the link function, which sets the scale on which effects are additive. Because a model without an interaction term is additive on the link scale, choosing a link is choosing what 'no interaction' means. The paper has two connected parts. First, a preregistered review of recent articles from five prominent psychology journals shows that interaction testing is routinely performed in research, that the link function is most often unspecified, and that constrained outcomes are frequently analyzed with Gaussian identity-link models. Second, worked examples and simulations examine forced-choice accuracy with a chance floor, the choice between logit and probit links, and sum scores as a measurement case, and quantify how a plausible-looking link can turn an additive structure on one scale into a statistically significant interaction on another. A final diagnostic simulation shows that residual checks, a link test, and AIC comparisons give uneven protection and do not by themselves decide which scale the scientific claim concerns. Interaction claims remain underspecified until researchers state, justify, and probe the scale on which no interaction is assumed."
#impact-statement: "This template helps you get started."
keywords: [interactions, moderation, generalized linear models, link functions, model misspecification]
bibliography: references.bib
floatsintext: true
format:
  # apaquarto-docx: default
  # apaquarto-html: default
  # apaquarto-typst: default
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    fig-pos: "h"
    header-includes:
      - \usepackage{float}
      - \raggedbottom
    include-in-header:
      text: |
        % Single-space table bodies so tables fit in text (floatsintext).
        % Wrapped lines inside a cell are set slightly tighter than normal and
        % each row gets a little extra vertical air, so a line break inside a
        % cell is not mistaken for the start of a new table row.
        \usepackage{etoolbox}
        \usepackage{array}
        \AtBeginEnvironment{longtable}{%
          \linespread{0.95}\selectfont
          \setlength{\extrarowheight}{6pt}}
        \AtBeginEnvironment{tabular}{%
          \linespread{0.95}\selectfont
          \setlength{\extrarowheight}{6pt}}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  dev = "pdf",
  fig.align = "center"
)
```

```{r}
#| label: packages
#| include: false

library(ggplot2)
library(stringr)
```

```{r}
#| label: ggplot-setup

# Theme copied locally so rendering has no dependency on deleted analysis helpers.
ggplot2::theme_set((ggplot2::theme_minimal(base_size = (10), base_family = "") +
    ggplot2::theme(
      plot.title = ggplot2::element_text(
        face = "bold",
        size = (10) + 1,
        margin = ggplot2::margin(b = 3)
      ),
      plot.subtitle = ggplot2::element_text(
        size = (10) - 1,
        color = "grey25",
        margin = ggplot2::margin(b = 6)
      ),
      axis.title = ggplot2::element_text(size = (10)),
      axis.text = ggplot2::element_text(size = (10) - 1, color = "grey20"),
      strip.text = ggplot2::element_text(face = "bold", size = (10) - 1),
      legend.position = "bottom",
      legend.title = ggplot2::element_text(size = (10) - 1),
      legend.text = ggplot2::element_text(size = (10) - 1),
      legend.key.width = grid::unit(1.25, "lines"),
      panel.grid.minor = ggplot2::element_blank(),
      panel.grid.major = ggplot2::element_line(linewidth = 0.25, color = "grey88"),
      panel.spacing = grid::unit(0.9, "lines"),
      plot.margin = ggplot2::margin(6, 8, 6, 8)
    )))
```

##

We start with a simplified example. Two groups of children, for instance a clinical and a neurotypical group, differ in how many errors they make on a task, and in both groups errors decrease with age. Imagine a researcher who records error counts from children aged six to ten and analyzes them, as is common, with a Gaussian linear model that includes an age-by-group interaction. In @fig-simple-count-example (Panel A) the two fitted lines decline at different rates, and the gap between the groups is much wider at age six than at age ten. Is there an age-by-group interaction?

```{r}
#| label: fig-simple-count-example
#| fig-cap: "Simulated error counts for two groups of children aged 6 to 10, generated with the same age effect in both groups on the log of the expected error count. A Gaussian linear model fitted to the counts produces straight lines with different slopes, which suggests an age-by-group interaction. A Poisson model with a log link fitted to the same data produces curves that converge with age on the count scale and parallel lines on the log scale, so the group gap shrinks on the count scale and stays constant on the log scale. Whether the pattern is an interaction depends on the scale on which effects are assumed to add."
#| fig.width: 7
#| fig.height: 3.2

set.seed(3)

N <- 250
d_count <- data.frame(
  age = round(runif(N, 6, 10), 1),
  condition = factor(rbinom(N, 1, .5), labels = c("Group A", "Group B"))
)

d_count$eta <- with(d_count, 2.6 - 0.55 * (age - 6) + 0.65 * (condition == "Group B"))
d_count$errors <- rpois(N, exp(d_count$eta))

fit_count <- glm(errors ~ age + condition, family = poisson(link = "log"), data = d_count)

count_pred <- expand.grid(age = seq(6, 10, length.out = 200), condition = levels(d_count$condition))
p <- predict(fit_count, newdata = count_pred, type = "link", se.fit = TRUE)

scale_levels <- c("Count scale: errors", "Log scale: log(errors)")

pred_two <- rbind(
  data.frame(
    count_pred, scale = scale_levels[1],
    fit = exp(p$fit),
    lower = exp(p$fit - 1.96 * p$se.fit),
    upper = exp(p$fit + 1.96 * p$se.fit)
  ),
  data.frame(
    count_pred, scale = scale_levels[2],
    fit = p$fit,
    lower = p$fit - 1.96 * p$se.fit,
    upper = p$fit + 1.96 * p$se.fit
  )
)
pred_two$scale <- factor(pred_two$scale, levels = scale_levels)

d_pos <- d_count[d_count$errors > 0, ]
points_two <- rbind(
  data.frame(age = d_count$age, condition = d_count$condition, y = d_count$errors, scale = scale_levels[1]),
  data.frame(age = d_pos$age, condition = d_pos$condition, y = log(d_pos$errors), scale = scale_levels[2])
)
points_two$scale <- factor(points_two$scale, levels = scale_levels)

ggplot(points_two, aes(age, y, color = condition, shape = condition)) +
  geom_point(size = 2.2, alpha = .5) +
  geom_ribbon(
    data = pred_two,
    aes(x = age, ymin = lower, ymax = upper, fill = condition, group = condition),
    inherit.aes = FALSE, alpha = .18, color = NA
  ) +
  geom_line(
    data = pred_two,
    aes(x = age, y = fit, linetype = condition, group = condition),
    linewidth = 1.1
  ) +
  facet_wrap(~ scale, scales = "free_y") +
  scale_color_manual(values = c("Group A" = "#0072B2", "Group B" = "#D55E00")) +
  scale_fill_manual(values = c("Group A" = "#0072B2", "Group B" = "#D55E00")) +
  labs(x = "Age", y = NULL, color = NULL, fill = NULL, shape = NULL, linetype = NULL) +
  (ggplot2::theme_minimal(base_size = (11), base_family = "") +
    ggplot2::theme(
      plot.title = ggplot2::element_text(
        face = "bold",
        size = (11) + 1,
        margin = ggplot2::margin(b = 3)
      ),
      plot.subtitle = ggplot2::element_text(
        size = (11) - 1,
        color = "grey25",
        margin = ggplot2::margin(b = 6)
      ),
      axis.title = ggplot2::element_text(size = (11)),
      axis.text = ggplot2::element_text(size = (11) - 1, color = "grey20"),
      strip.text = ggplot2::element_text(face = "bold", size = (11) - 1),
      legend.position = "bottom",
      legend.title = ggplot2::element_text(size = (11) - 1),
      legend.text = ggplot2::element_text(size = (11) - 1),
      legend.key.width = grid::unit(1.25, "lines"),
      panel.grid.minor = ggplot2::element_blank(),
      panel.grid.major = ggplot2::element_line(linewidth = 0.25, color = "grey88"),
      panel.spacing = grid::unit(0.9, "lines"),
      plot.margin = ggplot2::margin(6, 8, 6, 8)
    )) +
  theme(legend.position = "top")
```

An interaction is a difference between differences, so the question is whether the group difference itself changes with age. Since the gap narrows as the children grow older, the natural conclusion is that age and group interact. The data in @fig-simple-count-example, however, were generated without any interaction. Errors follow a Poisson process in which age and group add on the log scale, so the two groups share the same age slope on that scale. A Poisson model with a log link recovers this structure, and on the log scale its fitted lines are parallel (@fig-simple-count-example, Panel B). Mapped back onto the count scale, the same model produces two curves that converge with age, because a constant difference on the log scale is a constant ratio on the count scale, and the absolute gap between the groups therefore narrows as errors decline (@fig-simple-count-example, Panel C). Whether an interaction is present thus depends on the scale on which effects are assumed to add. Additivity on the log scale implies non-parallel curves on the count scale, and additivity on the count scale would imply non-parallel lines on the log scale.

This ambiguity extends far beyond this single illustration, as interaction hypotheses are widespread in psychological research and often represent the central target of inquiry [@rohrer2021; @domingue2024; @mccabe2022]. Whenever researchers ask whether an effect is moderated, or whether it holds only under certain conditions, they are asserting that the impact of one variable depends on another.

In psychology, this scale dependence has a characteristic source in measurement. Researchers routinely conceptualize constructs such as cognitive ability or symptom severity as continuous latent dimensions [@embretsonReise2000; @wagenmakers2012]. Those dimensions are rarely observed directly; what reaches the data are bounded, discrete quantities, such as counts, ratings, and accuracy scores that carry floors, ceilings, and chance levels. Each measurement scale imposes its own mapping from the latent dimension to the observed response. Where the mapping compresses, a constant effect on the latent dimension appears as a shrinking effect on the observed response: a group or condition closer to a bound has less room to move, and can look less affected by a manipulation, than a group or condition farther away from bounds (e.g., @fig-simple-count-example). 

Generalized linear models (GLMs) make this mapping explicit and give it a standard vocabulary [@nelderwedderburn1972; @mccullaghnelder1989], separating two choices that are easy to conflate, (1) the outcome family, which describes the response, and (2) the link function, which sets the scale on which effects combine (Box 1).

::: {.callout-note title="Box 1: Generalized linear models: outcome family and link function" icon="false"}
The *outcome family* is the distribution assumed for the response. It defines which values the outcome can take and how observations vary around their expectation. The *link function*, $g(\cdot)$, maps the expected outcome, $\mu_i = E(Y_i)$, onto the linear predictor, $\eta_i$, the scale on which the model is additive. With two predictors and their interaction term, $$g(\mu_i) = \eta_i, \qquad \eta_i = \beta_0 + \beta_1 X_i + \beta_2 Z_i + \beta_3 X_i Z_i.$$

The inverse link, $g^{-1}(\cdot)$, maps this additive structure back onto the expected response. Its familiar role is to keep expected values within the admissible range of a constrained outcome: $\exp(\eta)$ keeps expected counts positive, and the inverse logit keeps probabilities between 0 and 1. Less obviously, this also fixes what the absence of an interaction, $\beta_3 = 0$, means in the fitted model. The product-term coefficient $\beta_3$ is therefore an interaction on the link scale. A response-scale interaction is a different quantity: a difference between expected differences on the observed outcome, for example, whether the group difference in expected errors changes with age in @fig-simple-count-example. Unless the link is the identity, the two quantities can disagree [@ainorton2003; @mccabe2022; @mize2019]. Each family has a canonical link, which is the one that arises directly from the family's mathematical form, but it is a convenient default and other links are legitimate. A Gaussian linear model is the special case that combines a Gaussian family with an identity link, so the additive scale and the response scale coincide and the scale question is invisible. The model behind @fig-simple-count-example is a Poisson family with a log link. The family respects that errors are non-negative counts, and the link places additivity on the log scale.
:::

Often the outcome family is chosen with care while the link function is left at its software default, fixing the scale of additivity without deliberation. Two models can share a family and still place additivity on different scales. For example, a Gamma family suits reaction times data; within that family, a log link makes effects additive in log milliseconds, a proportional change, and the canonical inverse link makes them additive in reciprocal means, a change in speed. A binomial family fits correct-versus-incorrect responses and leaves the link open among logit, probit, chance-corrected, and other forms.

The scale dependence of interactions has long been recognized. @loftus1978 showed that interactions involving response probabilities are meaningful only relative to a mapping between observed and theoretical scales; @cox1984 formalized interaction as a model-relative concept tied to transformation and interpretation. @busemeyer1983 carried the same point into the measurement model, where the conclusion about how two variables combine depends on the model assumed to link responses to the quantities they measure. @lubinski1990 showed that the specification choices alone can produce the appearance of an interaction where the generating process contains none.

More recently, @wagenmakers2012 separated removable from nonremovable interactions. A removable interaction can be undone by a monotonic transformation of the dependent variable, whereas a nonremovable one cannot. In this regard, @dunnanderson2026 developed a removability test for linear models, while @rohrer2021 showed that the same interaction can be positive, null, or negative depending on the scale chosen for the outcome, so the appropriate test follows from the scale on which the hypothesis is posed. Rimpler and colleagues [-@rimpler2026] added that a product term assumes one predictor's effect changes at a constant rate across levels of the other, and so misrepresents any dependence that does not take that form, such as a threshold effect. Because the product-term coefficient of a GLM refers to the link scale (Box 1), a question about the response scale is answered with marginal effects [@ainorton2003; @mccabe2022; @mize2019; @rohrer2026]. 

Domingue and colleagues [-@domingue2024] demonstrated that outcome-model misspecification can produce false interaction findings for binary, count, and censored outcomes, and @liddellkruschke2018 showed that treating ordinal responses as metric can inflate error rates for interactions. Czado and Santner, and later Yu and Huang, studied the inferential consequences of link misspecification in binary GLMs and GLMMs [@czadosantner1992; @yuhuang2019]. 

Despite the above cited literature spanning decades, to our knowledge no study has attempted to quantify how often applied psychological research makes explicit the outcome family and link that define the no-interaction baseline. The present article approaches that gap from two directions. We first document current practice with a preregistered review of recent empirical articles in five prominent psychology journals, quantifying how often interactions are tested, how often the outcome family and link function are made explicit, and which response formats predominate. We then develop a framework that separates outcome family, link function, and target metric, and work it through three cases, forced-choice accuracy with a chance floor, the choice between logit and probit links, and sum scores. Simulations quantify how often an inadequate link turns a purely additive structure into a statistically significant interaction, and whether standard diagnostics detect the misspecification.

# Current Practice in Psychological Interaction Testing: A Preregistered Review

```{r review-section-load-data}
#| include: false

find_repo_file <- function(...) {
  rel <- file.path(...)
  candidates <- c(rel, file.path("..", rel))
  hit <- candidates[file.exists(candidates)]
  if (length(hit) == 0) {
    stop(
      "Cannot find required file: ", rel, ".\n",
      "Looked in: ", paste(candidates, collapse = "; "), ".\n",
      "Run the corresponding script before rendering paper.qmd."
    )
  }
  hit[1]
}

review_table_path <- find_repo_file("tables", "review-summary.csv")
outcome_table_path <- find_repo_file("tables", "review-outcome-types.csv")
agreement_table_path <- find_repo_file("tables", "review-intercoder-agreement.csv")
journal_table_path <- find_repo_file("tables", "review-by-journal.csv")
screening_flow_path <- find_repo_file("tables", "review-screening-flow.csv")

review_table <- read.csv(review_table_path)
outcome_table <- read.csv(outcome_table_path)
agreement_table <- read.csv(agreement_table_path)
journal_table <- read.csv(journal_table_path)
screening_flow_table <- read.csv(screening_flow_path)

get_review_row <- function(row_id, data = review_table) {
  if (!"row_id" %in% names(data)) {
    stop("Review summary table must contain a row_id column. Re-run scripts/01-review-descriptives.R before rendering.")
  }
  idx <- which(data$row_id == row_id)
  if (length(idx) != 1) {
    stop("Could not uniquely identify review row_id: ", row_id)
  }
  data[idx, ]
}

n_eligible <- get_review_row("eligible_empirical")
n_interactions <- get_review_row("testing_interactions")
n_non_identity <- get_review_row("non_identity_link")
n_explicit <- get_review_row("explicit_link")
n_incorrect_identity <- get_review_row("incorrect_identity")
n_significant_incorrect_identity <- get_review_row("significant_incorrect_identity")

get_outcome_row <- function(pattern, data = outcome_table) {
  idx <- grep(pattern, data$outcome_type, ignore.case = TRUE)
  if (length(idx) != 1) {
    stop("Could not uniquely identify outcome row: ", pattern)
  }
  data[idx, ]
}

get_agreement_row <- function(variable, data = agreement_table) {
  idx <- which(data$variable == variable)
  if (length(idx) != 1) {
    stop("Could not uniquely identify agreement variable: ", variable)
  }
  data[idx, ]
}

get_journal_row <- function(journal, data = journal_table) {
  idx <- which(data$journal == journal)
  if (length(idx) != 1) {
    stop("Could not uniquely identify journal row: ", journal)
  }
  data[idx, ]
}

outcome_binary <- get_outcome_row("Binary")
outcome_sum <- get_outcome_row("Sum scores")
outcome_ordinal <- get_outcome_row("Ordinal")
outcome_rt <- get_outcome_row("Response times")
outcome_neural <- get_outcome_row("Neural")
outcome_difference <- get_outcome_row("Difference")
outcome_correlations <- get_outcome_row("Correlations")
outcome_counts <- get_outcome_row("Counts")

agree_non_identity <- get_agreement_row("Uses_non_identity_link_function")
agree_explicit <- get_agreement_row("Explicit_link_function")
agree_incorrect_identity <- get_agreement_row("Incorrect_identity_link_function")
agree_significant <- get_agreement_row("Finds_significant_interaction")

journal_psychmed <- get_journal_row("Psychological Medicine")
journal_min_incorrect_identity <- journal_table[which.min(journal_table$incorrect_identity_percent_interaction), ]
journal_max_incorrect_identity <- journal_table[which.max(journal_table$incorrect_identity_percent_interaction), ]
journal_min_interaction <- journal_table[which.min(journal_table$interaction_testing_percent_of_eligible), ]
journal_max_interaction <- journal_table[which.max(journal_table$interaction_testing_percent_of_eligible), ]

sem_row_idx <- grep("SEM|latent-variable-only", screening_flow_table$step, ignore.case = TRUE)
if (length(sem_row_idx) != 1) {
  stop("Could not uniquely identify SEM / latent-variable-only exclusions row in screening_flow_table")
}
sem_exclusions_n <- screening_flow_table$n[[sem_row_idx]]

fmt_ci <- function(row) {
  paste0(row$percent, "%, 95% CI [", row$ci_low, ", ", row$ci_high, "]")
}

fmt_num <- function(x, digits = 3) {
  out <- ifelse(is.na(x), "--", formatC(as.numeric(x), format = "f", digits = digits))
  out
}

fmt_count <- function(x) {
  ifelse(is.na(x), "--", formatC(as.numeric(x), format = "f", digits = 0))
}

fmt_agreement_kappa <- function(row) {
  paste0(
    fmt_num(100 * row$agreement, 1), "% agreement, kappa = ",
    fmt_num(row$kappa, 2)
  )
}

fmt_yes_margins <- function(row) {
  if (all(c("coder1_yes_percent", "coder2_yes_percent") %in% names(row))) {
    return(paste0(
      fmt_num(row$coder1_yes_percent, 1), "% / ",
      fmt_num(row$coder2_yes_percent, 1), "%"
    ))
  }
  n <- row$n_double_coded
  coder1_yes <- 100 * (row$n_10 + row$n_11) / n
  coder2_yes <- 100 * (row$n_01 + row$n_11) / n
  paste0(fmt_num(coder1_yes, 1), "% / ", fmt_num(coder2_yes, 1), "%")
}

fmt_ci_cols <- function(low, high, digits = 3) {
  ifelse(
    is.na(low) | is.na(high),
    "--",
    paste0("[", fmt_num(low, digits), ", ", fmt_num(high, digits), "]")
  )
}

review_display_table <- function(dat) {
  out <- data.frame(
    Quantity = dat$quantity,
    n = fmt_count(dat$n),
    `Denom.` = fmt_count(dat$denominator_n),
    `%` = paste0(fmt_num(dat$percent, 1), "%"),
    `95% CI` = paste0("[", fmt_num(dat$ci_low, 1), "%, ", fmt_num(dat$ci_high, 1), "%]"),
    check.names = FALSE
  )
  out
}

simulation_display_table <- function(dat, effect_col = NULL, include_scale = FALSE) {
  out <- data.frame(.row = seq_len(nrow(dat)), stringsAsFactors = FALSE)
  if ("Scenario" %in% names(dat)) out$Scenario <- dat$Scenario
  if ("Generating link" %in% names(dat)) out$`Generating link` <- dat$`Generating link`
  if ("Model" %in% names(dat)) out$Model <- dat$Model
  if ("Fitted link" %in% names(dat)) out$`Fitted link` <- dat$`Fitted link`
  if ("Link status" %in% names(dat)) out$`Link status` <- dat$`Link status`
  if ("Rate type" %in% names(dat)) out$`Rate type` <- dat$`Rate type`
  if (include_scale && "Units for group-gap change" %in% names(dat)) {
    out$Scale <- ifelse(grepl("latent", dat$`Units for group-gap change`), "latent theta", "sum score")
  }
  if ("Successful fits" %in% names(dat)) out$Fits <- fmt_count(dat$`Successful fits`)
  if ("Rejections" %in% names(dat)) out$Reject <- fmt_count(dat$Rejections)
  if ("Rejection rate" %in% names(dat)) {
    out$`Rejection rate` <- fmt_num(dat$`Rejection rate`, 3)
  }
  if (all(c("CI low", "CI high") %in% names(dat))) {
    out$`95% CI` <- fmt_ci_cols(dat$`CI low`, dat$`CI high`, 3)
  }
  if (!is.null(effect_col) && effect_col %in% names(dat)) {
    out$`Median gap change` <- fmt_num(dat[[effect_col]], 3)
  }
  if ("Median interaction coefficient" %in% names(dat)) {
    out$`Median beta interaction` <- fmt_num(dat$`Median interaction coefficient`, 3)
  }
  if ("Median interaction SE" %in% names(dat)) {
    out$`Median SE` <- fmt_num(dat$`Median interaction SE`, 3)
  }
  out$.row <- NULL
  out
}

rate_display_table <- function(dat) {
  data.frame(
    Quantity = dat$Quantity,
    Fits = fmt_count(dat$`Successful fits`),
    Sig = fmt_count(dat$`Significant tests`),
    Rate = fmt_num(dat$Rate, 3),
    `95% CI` = fmt_ci_cols(dat$`CI low`, dat$`CI high`, 3),
    check.names = FALSE
  )
}

```

The review asks how often psychological articles test interactions, and how often they make the outcome family and link function explicit enough for a reader to know which interaction claim was tested. The coding is descriptive, so it documents reporting practice rather than the correctness of any substantive conclusion.

## Review Method

The protocol was preregistered before any record was screened (<https://osf.io/bdu73/>), together with the coding data dictionary. The essentials are given here; the full protocol, sampling frame, and reliability are in Supplement A.

Following the preregistration, we sampled one journal from each of five areas: *Psychological Science* (general), *Journal of Experimental Psychology: General* (experimental), *Journal of Personality and Social Psychology* (social), *Developmental Science* (developmental), and *Psychological Medicine* (clinical). The five were selected on preregistered discretionary criteria: a first-quartile 2023 Impact Factor, a broad rather than narrowly specialized scope, not primarily publishing reviews, and strong reputation in their field, as jointly judged by the coauthors. From each journal's full set of 2024 articles in Scopus, records were screened in a fixed random order until 40 eligible interaction-testing articles were reached, for a preregistered total of 200. Because screening stopped once 40 interaction-testing articles were reached, the number of records screened, and therefore the denominator of the eligibility proportion, varied across journals. Eligible articles were empirical reports of original research; reviews, meta-analyses, editorials, theoretical and qualitative-only articles, replication-only studies, and psychometric validation studies were excluded. Detailed coding was applied only to articles testing at least one interaction, and only to observed-outcome analyses: articles whose relevant tests were purely latent-variable or structural-equation models fell outside the coding scheme (`r sem_exclusions_n` such exclusions).

Each interaction-testing article was independently coded by two researchers, with disagreements resolved by discussion or by consulting the other coauthors. Generative AI tools were used only as an optional aid to locate passages within an article; all coding used in the review is human.

For each article, the coders recorded five things: whether at least one tested interaction used a non-identity link; whether the link function was clearly identifiable; whether at least one interaction applied a Gaussian-identity analysis to a constrained outcome; whether at least one such interaction was statistically significant, or interpreted analogously in a Bayesian analysis; and which response-variable types were analyzed. 

By a Gaussian-identity analysis on a constrained outcome we mean a Gaussian linear regression with an identity link applied to an outcome with formal constraints, such as bounds, discreteness, positivity, or a known chance level. This is a conservative screening flag, applied only when the constraint was identifiable from the report and does not imply the substantive conclusion is wrong. The label is defined relative to the constrained process that generates the outcome (if the observed score were explicitly justified as the target metric, identity-link additivity would instead be a defensible modeling choice, a distinction we develop in the following sections). 

Interrater agreement on the double-coded records was high (Cohen's $\kappa$ between `r fmt_num(min(agreement_table$kappa), 2)` and `r fmt_num(max(agreement_table$kappa), 2)`). Supplement A, Section S2, reports the preregistered variable definitions, the operational coding rules including borderline cases, per-variable agreement and coder margins, and the by-journal results.

## Review Findings

First, interaction testing was common. Screening yielded `r n_eligible$n` eligible empirical articles, of which  `r n_interactions$n`  tested at least one interaction (`r fmt_num(n_interactions$percent, 1)`%). Under the stopping rule this proportion describes the records screened rather than a fixed-denominator sample, so we report it descriptively and treat the per-journal proportions as the more informative summary: these ranged from  `r fmt_num(journal_min_interaction$interaction_testing_percent_of_eligible, 1)`% in  `r journal_min_interaction$journal`  to  `r fmt_num(journal_max_interaction$interaction_testing_percent_of_eligible, 1)`% in  `r journal_max_interaction$journal`. 

Second, model-scale reporting was uncommon. Among the `r n_interactions$n` interaction-testing articles, `r n_non_identity$n` used at least one non-identity link function (`r str_extract(fmt_ci(n_non_identity), "\\d+%")`), and only `r n_explicit$n` made the link function clearly identifiable (`r str_extract(fmt_ci(n_explicit),"\\d+%")`). The link function was therefore not clearly identifiable in `r fmt_num(100 - n_explicit$percent, 0)`% of interaction-testing articles.

Even the `r n_explicit$n` cases rest on a rule looser than the preregistered one, which required a declaration of the link and treated naming a family or a generalized linear model as insufficient. In coding we also accepted link-specific model names, because in practice they identify the link: "logistic regression" names the logit and "probit regression" the probit. We also accepted "Poisson regression", which names a family whose canonical link is the log. The stricter preregistered rule would give a lower count. 

Third, constrained outcomes were often analyzed on the identity scale. Among the `r n_interactions$n` interaction-testing articles, `r n_incorrect_identity$n` included at least one Gaussian-identity analysis on a constrained observed outcome (`r str_extract(fmt_ci(n_incorrect_identity),"\\d+%")`). Of these `r n_incorrect_identity$n` cases, `r n_significant_incorrect_identity$n` reported at least one statistically significant interaction (`r fmt_num(n_significant_incorrect_identity$percent, 0)`%). Because significance was coded only within the flagged analyses, this proportion describes those articles alone, and across all interaction-testing articles `r fmt_num(100 * n_significant_incorrect_identity$n / n_interactions$n, 0)`% combined a constrained outcome, an identity-link analysis, and a significant interaction claim. This is the same concern raised by work on outcome-model mismatch in interaction testing [@domingue2024; @liddellkruschke2018]. @tbl-current-practice summarizes these results.

```{r}
#| label: tbl-current-practice
#| tbl-cap: "Descriptive summary of the preregistered review. Gaussian-identity analyses on constrained outcomes are cases where the reported outcome had formal constraints that make identity-link additivity inappropriate as a generative model unless explicitly justified as an approximation or as the target observed-score metric. The interval on the eligibility row is reported for completeness only: screening stopped at 40 interaction-testing articles per journal, denominator was not fixed by design."
#| echo: false

knitr::kable(
  review_display_table(review_table),
  align = c("l", "r", "r", "r", "r")
)
```

The pattern was present across the sampled journals, where the proportion of interaction-testing articles with a Gaussian-identity analysis on a constrained observed outcome ranged from `r fmt_num(journal_min_incorrect_identity$incorrect_identity_percent_interaction, 0)`% in `r journal_min_incorrect_identity$journal` to `r fmt_num(journal_max_incorrect_identity$incorrect_identity_percent_interaction, 0)`% in `r journal_max_incorrect_identity$journal` (Supplement A). 

The outcomes used in interaction analyses were heterogeneous. The most common broad, non-exclusive categories were binary, accuracy, or proportion outcomes (`r outcome_binary$n` articles), sum scores or composites (`r outcome_sum$n` articles), ordinal, Likert, or rating outcomes (`r outcome_ordinal$n` articles), response times or durations (`r outcome_rt$n` articles), neural or physiological measures (`r outcome_neural$n` articles), difference or distance scores (`r outcome_difference$n` articles), counts or error counts (`r outcome_counts$n` articles), and correlations or associations (`r outcome_correlations$n` articles). These broad categories combine the more detailed labels used to classify the response variables during coding. The operational definitions and decision rules used to distinguish them are reported in Supplement A.


```{r}
#| label: tbl-review-outcome-types
#| tbl-cap: "Broad response-variable categories in interaction-testing articles. Categories are non-exclusive, so one article can contribute to more than one row."
#| echo: false

outcome_display <- data.frame(
  `Outcome type` = outcome_table$outcome_type,
  n = fmt_count(outcome_table$n),
  `Denom.` = fmt_count(outcome_table$denominator_n),
  `%` = paste0(fmt_num(outcome_table$percent, 1), "%"),
  check.names = FALSE
)
knitr::kable(outcome_display, align = c("l", "r", "r", "r"))
```

## What the Present Review Shows, What It Does Not

In these journals, then, interaction tests are common and the scale on which no interaction is assumed is rarely stated. The link function was clearly identifiable in only `r fmt_num(n_explicit$percent, 0)`% of interaction-testing articles, and `r fmt_num(n_incorrect_identity$percent, 0)`% applied an identity-link analysis to a constrained outcome. Bounded outcomes also dominated, with binary, accuracy, or proportion outcomes, sum scores, and ordinal ratings together appearing in 152 of the 200 articles (76%). For approximately continuous outcomes, additivity on the observed scale can be a defensible target, whereas for bounded and discrete outcomes it is a stronger assumption, and treating them as metric distorts estimates, error rates, and interaction conclusions [@liddellkruschke2018; @burknervuorre2019].

A descriptive review cannot establish how often published interactions are scale artifacts, because that judgment requires knowing the scale on which the data-generating process is additive. @dunnanderson2026 provide a direct test of removability, but it requires cell-level summaries and error estimates that the reviewed articles often did not report, and many designs used mixed-effects models, whose output does not reduce to cell-level means. The remainder of the paper therefore turns to simulations, where the generating process is known by construction, and focuses on the binomial and aggregated responses that dominated the corpus, including sums of binary or ordinal items.

# The Link Function Problem in Three Common Cases

To show what link misspecification does to interaction tests, we conducted three simulation studies. Each asks whether a data-generating process that is purely additive on one scale yields a statistically significant interaction when it is modeled on another scale.

1.  *Forced-Choice Accuracy*: This case holds the binomial outcome family fixed and evaluates whether the link function must explicitly encode a task-implied chance floor (e.g., a .50 guessing threshold).
2.  *Within-Family Link Choice (Logit vs. Probit)*: This case holds both the binary family and the 0-to-1 bounds fixed, testing whether choosing between these two standard link functions alters the results.
3.  *Sum Scores*: This case holds the latent generating process fixed and evaluates whether the bounded total shows an interaction that the latent scale does not.

All three take up the case in which the outcome family is already broadly plausible and the link alone sets a disputable no-interaction baseline, which is the common situation for bounded and chance-limited psychological outcomes. Work on misspecification has largely varied the outcome model as a whole [@domingue2024; @liddellkruschke2018], whereas here the family stays fixed and the link carries the entire disagreement about what counts as no interaction.

The simulations share one design. Data are generated with no product term on the generating scale, and models assuming additivity on that scale and on plausible alternative scales are fitted to the same data. The rate at which the matched model rejects a null product term is the *nominal rejection rate*. A product term estimated under a mismatched model is a genuine property of that model's scale [@loftus1978; @cox1984; @domingue2024; @mccabe2022], correct where it is computed and misleading only when the claim is carried to the generating scale. Because that scale is known in a simulation, we call such a result a *pseudo-interaction*, and the rate at which it reaches significance the *pseudo-interaction detection rate*.

The scenarios were chosen to represent plausible psychological data. Each reproduces a common interaction-testing situation, in which two well-established and large main effects, for example a group difference combined with a condition effect or a developmental trend, raise the question of whether one effect moderates the other. Most scenarios set these effects strong enough to push expected values close to a floor, ceiling, or chance level while keeping them inside the observed range, which is where link curvature acts most strongly, and scenarios away from the bounds serve as a comparison. Technical details are reported in Supplement A, and Supplement B reports a broader sensitivity atlas that varies sample size, main-effect magnitude, the intercept, and selected design parameters. A final simulation, in the next section, examines whether standard diagnostics identify the misspecified link.

```{r simulation-load-results}
#| include: false

read_required_table <- function(file, script_hint) {
  path <- find_repo_file("tables", file)
  dat <- read.csv(path, stringsAsFactors = FALSE)
  if (nrow(dat) == 0) {
    stop("The table ", path, " exists but has zero rows. Re-run ", script_hint, ".")
  }
  dat
}

select_required_columns <- function(dat, cols) {
  missing <- setdiff(cols, names(dat))
  if (length(missing) > 0) {
    stop(
      "The following required columns were not found: ",
      paste(missing, collapse = ", ")
    )
  }
  dat[, cols, drop = FALSE]
}

rename_columns <- function(dat, labels) {
  names(dat)[match(names(labels), names(dat))] <- unname(labels)
  dat
}

# Inline row lookups must hit exactly one row. A label that no longer matches the
# summary tables would otherwise return a zero-length value that knitr renders as
# an empty string, silently dropping the number from the text.
require_single <- function(value, what) {
  if (length(value) != 1) {
    stop(
      "Inline lookup ", what, " matched ", length(value),
      " values instead of 1. Check that the label matches the summary table."
    )
  }
  value
}

forced_choice_sim_raw <- read_required_table(
  "simulation-summary-forced-choice.csv",
  "scripts/03a-simulation-forced-choice.R"
)
within_family_sim_raw <- read_required_table(
  "simulation-summary-within-family-links.csv",
  "scripts/03b-simulation-within-family-links.R"
)
sum_score_sim_raw <- read_required_table(
  "simulation-summary-sum-scores.csv",
  "scripts/03c-simulation-sum-scores.R"
)
diagnostic_sim_raw <- read_required_table(
  "simulation-summary-diagnostic-worked-example.csv",
  "scripts/04-diagnostic-worked-example.R"
)

diagnostic_scenarios_raw <- read_required_table(
  "simulation-summary-diagnostic-scenarios.csv",
  "scripts/04-diagnostic-worked-example.R"
)

diagnostic_calibration_raw <- read_required_table(
  "diagnostic-calibration.csv",
  "scripts/04-diagnostic-worked-example.R"
)

diagnostic_replications_path <- find_repo_file(
  "outputs", "diagnostic-simulation-replications.csv"
)
diagnostic_replications <- read.csv(
  diagnostic_replications_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

diagnostic_dharma_cols <- c(
  "dharma_uniformity_p",
  "dharma_dispersion_p",
  "dharma_quantile_fitted_p",
  "dharma_quantile_predictor_p",
  "dharma_categorical_design_p"
)
diagnostic_required_cols <- c(
  "scenario", "interaction_p", diagnostic_dharma_cols,
  "aic_fitted_minus_true"
)
if (!all(diagnostic_required_cols %in% names(diagnostic_replications))) {
  stop(
    "The diagnostic replication file is missing required columns: ",
    paste(
      setdiff(diagnostic_required_cols, names(diagnostic_replications)),
      collapse = ", "
    )
  )
}

diagnostic_calibration_required_cols <- c(
  "scenario", "diagnostic", "model_specification", "n_attempted",
  "n_successful", "flagging_rate", "ci_low", "ci_high"
)
if (!all(diagnostic_calibration_required_cols %in% names(diagnostic_calibration_raw))) {
  stop(
    "The diagnostic calibration table is missing required columns: ",
    paste(
      setdiff(
        diagnostic_calibration_required_cols,
        names(diagnostic_calibration_raw)
      ),
      collapse = ", "
    )
  )
}
diagnostic_calibration_correct <- diagnostic_calibration_raw[
  diagnostic_calibration_raw$model_specification == "Correct link" &
    diagnostic_calibration_raw$n_successful > 0 &
    is.finite(diagnostic_calibration_raw$flagging_rate),
  ,
  drop = FALSE
]
# In the saturated 2 x 2 interaction model the squared population-level linear
# predictor is collinear with the existing fixed effects, so the
# Pregibon-style check is not applicable. The continuous-predictor GLMM remains
# estimable and is calibrated separately below.
pregibon_na_scenario <- "Binary repeated trials"
pregibon_glmm_continuous_scenario <- "Binary continuous predictor"

diagnostic_calibration_dharma <- diagnostic_calibration_correct[
  grepl("^DHARMa", diagnostic_calibration_correct$diagnostic), , drop = FALSE
]
diagnostic_calibration_pregibon <- diagnostic_calibration_correct[
  grepl("Pregibon", diagnostic_calibration_correct$diagnostic), , drop = FALSE
]
dharma_calibration_max <- 100 * max(
  diagnostic_calibration_dharma$flagging_rate,
  na.rm = TRUE
)
pregibon_calibration_single_level_max <- 100 * max(
  diagnostic_calibration_pregibon$flagging_rate[
    !(diagnostic_calibration_pregibon$scenario %in% c(
      pregibon_na_scenario,
      pregibon_glmm_continuous_scenario
    ))
  ],
  na.rm = TRUE
)
pregibon_calibration_glmm_continuous <- 100 * require_single(
  diagnostic_calibration_pregibon$flagging_rate[
    diagnostic_calibration_pregibon$scenario == pregibon_glmm_continuous_scenario
  ],
  "correct-link Pregibon calibration for the continuous-predictor GLMM"
)

diagnostic_calibration_regular <- (
  nrow(diagnostic_calibration_correct) > 0 &&
    all(diagnostic_calibration_correct$ci_low <= .05, na.rm = TRUE)
)
diagnostic_calibration_result_sentence <- if (diagnostic_calibration_regular) {
  paste(
    "Baseline flagging rates under the correct link were generally low",
    "across diagnostics and scenarios; full calibration results are reported",
    "in Supplement A"
  )
} else {
  paste(
    "Calibration rates varied across diagnostics or scenarios, so flagging",
    "rates under misspecification should be read against the baseline rates",
    "reported in Supplement A."
  )
}

# Scenario tables are produced by the revised scripts and document the
# deterministic values implied by each data-generating scenario. They are
# loaded here so that the manuscript and scripts remain synchronized, but
# they are not printed as extra main-text tables. The multi-panel figures
# are the primary scenario visualizations.
forced_choice_scenario <- read_required_table(
  "scenario-table-forced-choice.csv",
  "scripts/03a-simulation-forced-choice.R"
)
within_family_scenario <- read_required_table(
  "scenario-table-within-family-links.csv",
  "scripts/03b-simulation-within-family-links.R"
)
sum_score_scenario <- read_required_table(
  "scenario-table-sum-scores.csv",
  "scripts/03c-simulation-sum-scores.R"
)
diagnostic_scenario <- read_required_table(
  "scenario-table-diagnostic-worked-example.csv",
  "scripts/04-diagnostic-worked-example.R"
)

forced_choice_sim <- select_required_columns(
  forced_choice_sim_raw,
  c(
    "scenario",
    "model",
    "n_total",
    "n_fit_ok",
    "n_fit_problem",
    "fit_problem_rate",
    "n_p_finite_ok",
    "n_rejections_ok",
    "rejection_rate_ok",
    "rate_type",
    "ci_low_ok",
    "ci_high_ok",
    "n_p_finite_problem",
    "n_rejections_problem",
    "rejection_rate_problem"
  )
)
forced_choice_sim <- rename_columns(
  forced_choice_sim,
  c(
    scenario = "Scenario",
    model = "Model",
    n_total = "Fits attempted",
    n_fit_ok = "Fits passing checks",
    n_fit_problem = "Flagged fits",
    fit_problem_rate = "Fit-problem rate",
    n_p_finite_ok = "Finite p-values, acceptable fits",
    n_rejections_ok = "Rejections, acceptable fits",
    rejection_rate_ok = "Primary rejection rate",
    rate_type = "Rate type",
    ci_low_ok = "Primary CI low",
    ci_high_ok = "Primary CI high",
    n_p_finite_problem = "Finite p-values, flagged fits",
    n_rejections_problem = "Apparent rejections, flagged fits",
    rejection_rate_problem = "Apparent rejection rate, flagged fits"
  )
)

sum_score_sim <- select_required_columns(
  sum_score_sim_raw,
  c(
    "scenario",
    "model",
    "change_in_group_difference_units",
    "n_successful_fits",
    "n_rejections",
    "rejection_rate",
    "rate_type",
    "ci_low",
    "ci_high",
    "median_change_in_group_difference_outcome_units"
  )
)
sum_score_sim <- rename_columns(
  sum_score_sim,
  c(
    scenario = "Scenario",
    model = "Model",
    change_in_group_difference_units = "Units for group-gap change",
    n_successful_fits = "Successful fits",
    n_rejections = "Rejections",
    rejection_rate = "Rejection rate",
    rate_type = "Rate type",
    ci_low = "CI low",
    ci_high = "CI high",
    median_change_in_group_difference_outcome_units = "Median change in group gap"
  )
)

# The within-family script writes repeated-trial GLMM results without a
# scenario column or response-scale gap-change summaries.
within_family_sim <- select_required_columns(
  within_family_sim_raw,
  c(
    "generating_link",
    "fitted_link",
    "link_match",
    "n_successful_fits",
    "n_rejections",
    "rejection_rate",
    "rate_type",
    "ci_low",
    "ci_high",
    "median_interaction_coef",
    "median_interaction_se",
    "deterministic_product_on_fitted_link_scale"
  )
)
within_family_sim <- rename_columns(
  within_family_sim,
  c(
    generating_link = "Generating link",
    fitted_link = "Fitted link",
    link_match = "Link status",
    n_successful_fits = "Successful fits",
    n_rejections = "Rejections",
    rejection_rate = "Rejection rate",
    rate_type = "Rate type",
    ci_low = "CI low",
    ci_high = "CI high",
    median_interaction_coef = "Median interaction coefficient",
    median_interaction_se = "Median interaction SE",
    deterministic_product_on_fitted_link_scale = "Induced product term"
  )
)
diagnostic_sim <- select_required_columns(
  diagnostic_sim_raw,
  c("scenario", "quantity", "n_successful_fits", "n_significant", "rate", "ci_low", "ci_high")
)

diagnostic_sim <- rename_columns(
  diagnostic_sim,
  c(
    scenario = "Scenario",
    quantity = "Quantity",
    n_successful_fits = "Successful fits",
    n_significant = "Significant tests",
    rate = "Rate",
    ci_low = "CI low",
    ci_high = "CI high"
  )
)

diagnostic_scenarios <- select_required_columns(
  diagnostic_scenarios_raw,
  c(
    "scenario",
    "paper_anchor",
    "outcome_family",
    "true_model_label",
    "fitted_model_label",
    "n_replications",
    "pseudo_interaction_detection_rate",
    "pseudo_interaction_detection_ci_low",
    "pseudo_interaction_detection_ci_high",
    "dharma_detection_rate_min",
    "dharma_detection_rate_max",
    "dharma_n_checks_used",
    "pregibon_detection_rate",
    "aic_favors_target",
    "quantitative_description",
    "variability_summary"
  )
)

diagnostic_scenarios <- rename_columns(
  diagnostic_scenarios,
  c(
    scenario = "Case",
    paper_anchor = "Paper anchor",
    outcome_family = "Outcome family",
    true_model_label = "Target model",
    fitted_model_label = "Misspecified model",
    n_replications = "Replications",
    pseudo_interaction_detection_rate = "Pseudo-interaction detection",
    pseudo_interaction_detection_ci_low = "Pseudo-interaction CI low",
    pseudo_interaction_detection_ci_high = "Pseudo-interaction CI high",
    dharma_detection_rate_min = "DHARMa min",
    dharma_detection_rate_max = "DHARMa max",
    dharma_n_checks_used = "DHARMa n checks",
    pregibon_detection_rate = "Pregibon check",
    aic_favors_target = "AIC favors target",
    quantitative_description = "Quantitative description",
    variability_summary = "Variability summary"
  )
)

diagnostic_scenario_display_table <- function(dat) {
  data.frame(
    Case = dat$Case,
    `Target model` = dat$`Target model`,
    `Misspecified model` = dat$`Misspecified model`,
    `Pseudo-interaction rate` = paste0(
      fmt_num(dat$`Pseudo-interaction detection`, 3), " ",
      fmt_ci_cols(dat$`Pseudo-interaction CI low`, dat$`Pseudo-interaction CI high`, 3)
    ),
    DHARMa = paste0(
      fmt_num(dat$`DHARMa min`, 3), "-",
      fmt_num(dat$`DHARMa max`, 3),
      " (", dat$`DHARMa n checks`, ")"
    ),
    Pregibon = ifelse(
      dat$Case == pregibon_na_scenario,
      "N.A.",
      fmt_num(dat$`Pregibon check`, 3)
    ),
    `AIC favors target` = fmt_num(dat$`AIC favors target`, 3),
    check.names = FALSE
  )
}

dg_lookup <- function(case, col) {
  if (!col %in% names(diagnostic_scenarios)) {
    stop(
      "Unknown diagnostic scenario column: ", col,
      ". Available columns: ",
      paste(names(diagnostic_scenarios), collapse = ", ")
    )
  }
  require_single(
    diagnostic_scenarios[diagnostic_scenarios$Case == case, col, drop = TRUE],
    paste0("diagnostic_scenarios[", case, " / ", col, "]")
  )
}
dg_pct <- function(case, col) round(100 * dg_lookup(case, col))
dg_n <- function(case) dg_lookup(case, "Replications")
dg_max_other <- function(col, exclude = "Count errors") {
  round(100 * max(diagnostic_scenarios[diagnostic_scenarios$Case != exclude, col], na.rm = TRUE))
}
dg_range <- function(col) {
  v <- 100 * diagnostic_scenarios[[col]]
  sprintf("%d%% to %d%%", round(min(v, na.rm = TRUE)), round(max(v, na.rm = TRUE)))
}

# Replications attempted per diagnostic scenario, before any convergence loss.
n_rep_diagnostic <- max(table(diagnostic_replications$scenario))

# Participants in the binary diagnostic scenarios (see scripts/04-diagnostic-worked-example.R).
dg_N_binary <- 300
dg_val <- function(case, col, digits = 1) fmt_num(dg_lookup(case, col), digits)
```

## Forced-Choice Accuracy and the Non-Zero Chance Floor

If a participant answers $Y_i$ of $k$ trials correctly, a binomial sampling model, $Y_i \sim \mathrm{Binomial}(k, p_i)$, is the natural starting point. The family requires the expected accuracy to lie between 0 and 1. When guessing is a plausible process for the task, the range that matters is narrower, since a participant who knows nothing still gets a share $c$ of the trials right, $c = .50$ in a two-alternative or true/false task and $c = 1/m$ with $m$ alternatives, so expected accuracy lies between $c$ and 1, even though observed proportions can fall below $c$ through sampling variation. Accuracy data of this kind are analyzed in several ways in practice, on the raw proportion scale with a Gaussian identity model, with a standard binomial logit or probit, and, less often, with a link that encodes the chance floor.

These specifications differ in which of the two ranges they impose. The Gaussian identity model imposes neither, since it ignores the trial process, assumes additivity in percentage-correct units, and can predict outside the 0-to-1 range. The other two keep the binomial family and differ only in the link. The standard logit or probit, $p_i = F(\eta_i)$ with $F$ the logistic or standard-normal CDF, holds expected accuracy between 0 and 1 and places the lower asymptote at 0, whereas the chance-corrected link maps the linear predictor onto the narrower interval from $c$ to 1,
$p_i = c + (1 - c)F(\eta_i)$, which for $c = .50$ becomes $p_i = .50 + .50\,F(\eta_i)$. An interaction term in the standard model therefore tests additivity over the full 0-to-1 range, whereas the same term in the chance-corrected model tests additivity on a performance-above-chance scale, so the same four cell probabilities can imply different interactions on the three scales (@fig-motivating-example).

![A minimal $2 \times 2$ design crosses a condition (x-axis) with a group (color), shown on three scales. The four cell probabilities are identical in every panel; only the scale on which the interaction is evaluated changes. The data were generated with no condition-by-group interaction on the chance-corrected logit scale, which has a $.50$ chance floor, so the interaction coefficient is exactly zero on that scale. The same cells imply a positive interaction on the standard logit scale and a negative interaction on the raw probability scale. Light horizontal lines are equally spaced on each panel's own link scale, so equal steps on that scale map to unequal steps on the probability axis.](../figs/motivating-example.pdf){#fig-motivating-example}


The chance-corrected link is most defensible when the chance level is known from the task design, guessing is a plausible lower-bound process, and below-chance performance is not substantively meaningful for the target claim. In the simplest case, the chance level can be fixed by design, for example $c=0.50$ in a two-alternative task. In other cases, analysts may estimate the lower asymptote or report sensitivity analyses over plausible values of $c$. Lapses, response biases, item heterogeneity, changes across trials, or systematic below-chance performance may instead require richer psychometric solutions such as item-response models [e.g., @embretsonReise2000; @wichmann2001; @knoblauch2012]. For example, the 3-parameter logistic Item Response Theory model estimates a guessing parameter for each item, where the chance-corrected link fixes a single chance level for the whole design [@embretsonReise2000]. The chance-corrected link is thus a design-level approximation, and item-level measurement models remain the appropriate tool when the data and the scientific question call for that additional structure.

### Simulation 1: Design and Results

```{r}
#| label: forced-choice-inline
#| include: false

# Design constants for Simulation 1 (see scripts/03a-simulation-forced-choice.R).
sim_N <- 250
sim_k <- 20
settings_age_low <- 6
settings_age_high <- 10
n_rep <- max(forced_choice_sim$`Fits attempted`)
```

In each of `r n_rep` replications we generated 20 trial-level Bernoulli responses for each of $N = `r sim_N`$ participants. Age and group were constant across a participant's trials. The data-generating model included a subject random intercept with standard deviation $\sqrt{.30(\pi^2/3)/(1-.30)} \approx 1.187$, corresponding to latent-scale ICC $.30$ under the usual logistic convention. The probability of a correct response followed a chance-corrected logit with a .50 floor, a positive age effect, a group difference, and no age-by-group product term on the conditional chance-corrected-logit scale. Three scenarios differ only in overall performance, set by shifting the fixed intercept. The deterministic curves in Panel A are evaluated at random intercept 0.

Each simulated dataset was analyzed with four mixed models sharing the same subject random intercept: a Gaussian identity LMM applied to the binary trial outcome, standard binomial-logit and binomial-probit GLMMs, and the matched chance-corrected binomial-logit GLMM. The matched model was fitted with `lme4::glmer` and `psyphy::mafc.logit(2)`. Each model received one ordinary fit attempt. The primary rate uses finite Wald-normal product-term tests from fits without an lme4 convergence message, singularity, targeted convergence or Hessian warning during inference extraction, or non-finite inferential quantity. Finite $p$-values from flagged fits are retained and summarized separately in Supplement A.

![Simulation 1: forced-choice accuracy with a chance floor and a subject random intercept. The product term is zero on the conditional chance-corrected-logit scale. Panel A shows fixed-effect curves at random intercept 0 across three performance levels; the grey region lies below the .50 chance floor. Panel B shows the primary product-term rejection rates among fits passing the minimal fit-quality checks. The dashed line marks alpha = .05.](../figs/forced-choice-simulation.pdf){#fig-forced-choice-simulation fig-pos="H"}

For the matched chance-corrected model, the primary rate is the nominal rejection rate among numerically acceptable fits. For the three alternative links it is the pseudo-interaction detection rate among acceptable fits. Rates among flagged fits are descriptive apparent rejection rates, not valid inferential error rates. Numerical results in this section and Supplement A require regeneration with the revised mixed-model simulation.

## Within-Family Link Choice

In the forced-choice case the standard link was wrong for the task by construction, because it placed the lower asymptote at 0 when the design fixed it at chance. The link-function problem can also arise in less obvious cases. Logit and probit are both standard for binary data, respect the same 0-to-1 bounds, and usually produce very similar fitted probabilities. They differ mainly in interpretation, since logit coefficients are log-odds, whereas probit coefficients are expressed on a standard-normal latent scale, which makes probit the usual choice when modeling psychological quantities such as ability, signal strength, and response tendency [@dunn2018]. The choice between them is therefore often guided by convention, interpretation, or software, with little expected impact on substantive conclusions [@yuhuang2019; @agresti2002]. Yet even two nearly interchangeable links define slightly different scales of additivity. Equal differences under one link are not exactly equal under the other, especially toward the tails, and this alone can alter the estimated product term.

### Simulation 2: Design and Results

```{r}
#| label: within-family-inline
#| include: false

# Design constants for Simulation 2 (see scripts/03b-simulation-within-family-links.R).
wf_N <- 700
wf_k <- 15
n_rep_wf <- max(within_family_sim$`Successful fits`)

wf_lookup <- function(gen, fit, col) {
  require_single(
    within_family_sim[
      within_family_sim$`Generating link` == gen & within_family_sim$`Fitted link` == fit,
      col, drop = TRUE
    ],
    paste0("within_family_sim[", gen, " / ", fit, " / ", col, "]")
  )
}
wf_rejection_rate <- function(gen, fit) round(100 * wf_lookup(gen, fit, "Rejection rate"))
wf_induced_product <- function(gen, fit) fmt_num(wf_lookup(gen, fit, "Induced product term"), 3)
```

In each of `r n_rep_wf` replications we generated repeated binary responses for $N = `r wf_N`$ participants in a two-group, two-condition design, with `r wf_k` trials per participant-condition cell and a subject random intercept (latent intraclass correlation 0.30). The large sample was chosen so that the small systematic discrepancy between logit and probit links would be detectable, and it is larger than in a typical psychological study. The condition-by-group interaction term was zero on the generating link scale. We crossed the generating link with the fitted link, using logit and probit in both roles, and fitted every model as a random-intercept GLMM, recording how often each returned a significant condition-by-group interaction term at $\alpha = .05$. @fig-within-family-links shows the results; the full crossed-link rates are tabulated in Supplement A. 

![Simulation 2: logit versus probit in repeated-trial random-intercept GLMMs. Data are generated with no condition-by-group product term on the generating link scale. Panel A shows the corresponding cell probabilities under logit and probit links. Panel B shows product-term rejection rates for each generating–fitted link combination. Matched links remain near nominal alpha, whereas wrong links produce pseudo-interactions. Error bars are Wilson 95% confidence intervals; the dashed line marks alpha = .05.](../figs/within-family-links.pdf){#fig-within-family-links fig-pos="H"}

When the fitted link matches the generating link, the nominal rejection rate stays near .05, at `r wf_rejection_rate("logit", "logit")`% for logit fitted to logit-generated data and `r wf_rejection_rate("probit", "probit")`% for probit fitted to probit-generated data. Crossing the links raises the pseudo-interaction detection rate to `r wf_rejection_rate("probit", "logit")`% when a logit model is fitted to probit-generated data and `r wf_rejection_rate("logit", "probit")`% when a probit model is fitted to logit-generated data, roughly three times the nominal rate. The wrong-link detection rates are smaller than in the forced-choice case, and they appear among links that share the same family, the same 0-to-1 bounds, and nearly overlapping fitted curves in the middle of the probability range, the setting many applied analysts treat as interchangeable.

These rates reflect a fixed quantity. Under the wrong link the same cell probabilities imply a small but deterministic nonzero product term on the fitted scale. Re-expressing the cell probabilities of each generating model on the other link scale gives a product term of `r wf_induced_product("logit", "probit")` for logit-generated data fitted with a probit and `r wf_induced_product("probit", "logit")` for probit-generated data fitted with a logit, against exactly zero whenever the links match (Supplement A, Section S5), so the rejection rates above are the power to detect these values. Supplement B varies the intercept, which shifts where the four cell probabilities sit between 0 and 1, and shows that detectability changes non-monotonically with it, so the same link mismatch is nearly undetectable at some probability levels and much easier to detect at others. 

## Sum Scores

Many psychological outcomes are computed as total sum scores across questionnaire items, task items, symptoms, errors, or ratings. In the two cases above the outcome was a directly observed response, and only the link was in question. A sum score is different, because the score itself is already the result of an analytic choice. What is observed are the individual item responses, which are binary or ordinal; the total is a summary of them, produced by a scoring rule that gives every item the same weight and every point of the scale the same width. In psychological research the construct of interest is often the underlying, generally continuous dimension, so the total can be seen as a remapped version of that continuum. The link question then arrives together with a prior question about which metric the claim is about.

The bounds matter because the construct may have no floor or ceiling, whereas the total score has both. The relation between the two can be described by how much change on the latent dimension is needed to raise the total by one point, and this amount is not constant. In the middle of the score range a small latent change moves the total by a point, because many items are close to changing category. Near either bound a much larger latent change is needed, because most items are already at their lowest or highest category. A constant latent effect therefore translates into fewer raw-score points as the total approaches the floor or the ceiling, and the same holds wherever the instrument is less informative along the latent dimension, so the effect is not limited to the endpoints. A Gaussian identity model fitted to the total assumes the opposite, namely that a predictor shifts the score by the same number of points at every level of the other predictor. When one group lies closer to a bound, a predictor with the same latent effect in both groups moves that group's score by fewer points, and the model reports the difference as an interaction.

The problem we raise sits next to a recent, wider debate about whether sum scores should be replaced by factor scores or latent-variable models. Some argue that sum scores are transparent, reliable enough for many purposes, comparable across studies, and tied to how instruments are scored in applied settings [@widaman2023]; the counterargument is that a total encodes strong and usually untested assumptions about how items represent the construct [@mcneish2020; @mcneish2023]. Without entering that debate, we add a narrower point that it leaves aside. A model of the manifest total defines the absence of an interaction in raw-score points, so an interaction found there is an interaction on the score, and it becomes a claim about the construct only under the additional assumption that the two are linearly related.

When item-level data are available and the claim concerns latent constructs, an explicit measurement model such as SEM, item-level ordinal regression, or IRT-inspired specifications could be preferred, since it addresses the mapping directly, though at the price of moving the assumptions into the measurement model [@burknervuorre2019]. Often, however, researchers deliberately work with the total score for the practical reasons listed above. One option is then a family that respects the bounds and the discreteness of the total, for example a beta-binomial likelihood for a bounded count [@wilcox1981]. Here we use a deliberately unusual alternative that changes only the link, keeping the observed total under a Gaussian likelihood, rescaling it to the proportion of the maximum score, and applying a probit link, as in Simulation 3 below. This is an *ad hoc* device outside standard practice, and it illustrates whether the interaction conclusion changes if additivity is placed on a nonlinear scale that compresses the expected values toward the score bounds. It remains a model of the total, it will not recover the construct, and the Gaussian likelihood is still misspecified in its variance, since a bounded score has a variance that depends on its mean, so the probit link changes only the scale on which additivity is defined and leaves the distributional form as it is. Read against the sum-score debate, it is a precaution for researchers who have decided to analyze totals anyway and who accept that the interaction still needs a scale on which additivity is plausible.

### Simulation 3: Design and Results

```{r}
#| label: sum-score-inline
#| include: false

# Design constants for Simulation 3 (see scripts/03c-simulation-sum-scores.R).
ss_N   <- 600
ss_J   <- 9
ss_cat <- 3                 # each item scored 0 to 3
ss_max <- ss_J * ss_cat     # 0-27 total
n_rep_ss <- max(sum_score_sim$`Successful fits`)

ss_lookup <- function(scn, mdl, col) {
  require_single(
    sum_score_sim[
      sum_score_sim$Scenario == scn & sum_score_sim$Model == mdl,
      col, drop = TRUE
    ],
    paste0("sum_score_sim[", scn, " / ", mdl, " / ", col, "]")
  )
}
ss_rejection_rate <- function(scn, mdl) round(100 * ss_lookup(scn, mdl, "Rejection rate"))

# Ranges quoted in the text are the observed min/max across scenarios, so they
# stay ordered if a re-run changes which scenario is the extreme one.
ss_rate_range <- function(mdl) {
  rates <- sum_score_sim$`Rejection rate`[sum_score_sim$Model == mdl]
  if (length(rates) == 0) stop("No rejection rates found for model ", mdl, ".")
  round(100 * range(rates))
}
ss_rate_min <- function(mdl) ss_rate_range(mdl)[1]
ss_rate_max <- function(mdl) ss_rate_range(mdl)[2]
```

In each of `r n_rep_ss` replications, we generated data for $N = `r ss_N`$ participants from a continuous latent variable with a continuous predictor $x$, a group effect, and no x-by-group interaction. The latent values generated `r ss_J` ordinal items scored 0 to `r ss_cat`, which were summed to a total ranging from 0 to `r ss_max`. Item thresholds were shifted across three scenarios so that expected totals fell in the lower, middle, or upper part of the score range.

We analyzed each total in three ways: a Gaussian identity model of the observed sum score, a Gaussian-probit bounded-score model, and the true latent generating scale. The bounded-score model applies a probit link to a continuity-corrected score proportion and rescales predictions to the original sum-score metric. We used probit because its standard-normal linear predictor provides a simple correspondence with the latent scale. We recorded the proportion of significant x-by-group interactions at $\alpha = .05$. Results are shown in @fig-sum-score-simulation, with per-scenario rates in Supplement A, Section S4.

![Simulation 3: sum scores as bounded and discrete outcomes. Data are generated with no x-by-group product term on the latent scale. Panel A shows how the same additive latent structure maps onto sum scores in the lower, middle, and upper parts of the score range. Panel B shows product-term rejection rates for the latent benchmark and two observed-score models. The latent model remains near nominal alpha, whereas the observed-score models can produce pseudo-interactions. The dashed line marks alpha = .05. The latent benchmark is available only in simulation.](../figs/sum-score-simulation.pdf){#fig-sum-score-simulation fig-pos="H"}

The latent benchmark remained close to the nominal .05 rejection rate in every scenario, from `r ss_rate_min("Latent generating scale")`% to `r ss_rate_max("Latent generating scale")`%. The identity model rejected in `r ss_rejection_rate("Lower range", "Observed sum-score identity")`% of replications in the lower range, `r ss_rejection_rate("Middle range", "Observed sum-score identity")`% in the middle range, and `r ss_rejection_rate("Upper range", "Observed sum-score identity")`% in the upper range, so it came closest to the nominal rate in the middle, where the sum score is approximately linear in the latent variable, and its rate rose as expected totals approached a bound. Compression of the latent scale makes the same latent group difference appear to change across levels of the predictor. The bounded-score probit model rejected in `r ss_rejection_rate("Lower range", "Gaussian-probit bounded score")`%, `r ss_rejection_rate("Middle range", "Gaussian-probit bounded score")`%, and `r ss_rejection_rate("Upper range", "Gaussian-probit bounded score")`% of replications in the same three ranges. It respects the score bounds, although it is still a model of the manifest total.

This pattern mainly depends on where observations fall within the score range. Across the broader scenarios in Supplement B, the position of expected scores within the range affected identity-model rejection much more than sample size, item count, or latent dispersion. Raw-score additivity can approximate latent-scale additivity when scores remain away from the bounds. The bounded-score probit model can therefore serve as a sensitivity analysis. For claims about interactions between latent constructs, an explicit measurement model provides a more direct approach, for example SEM, item-level ordinal models, or IRT-inspired models. These approaches still depend on assumptions about thresholds, links, loadings, and latent-response scales, but they make those assumptions part of the interaction model.

# Can Standard Diagnostics Catch a Wrong Link?

The simulations above assume that the analyst accepts the fitted model. In practice, a researcher who finds an interaction may check the model before reporting it. We examine three checks that follow different strategies. A Pregibon-style link test probes the assumed link, residual diagnostics screen for misfit of any kind, and AIC compares candidate models. 

The Pregibon test refits the model with the squared fitted linear predictor, $\hat{\eta}^2$, as an additional predictor [@pregibon1980]. Under an adequate link this term carries no further information, and a significant coefficient indicates curvature that the link leaves unexplained.

Residual diagnostics locate each observed value within the distribution the model predicts for it. That position lies between 0 and 1 and is uniform under an adequate model whatever the family, so misfit appears as departure from uniformity, as over- or underdispersion, or as structure over fitted values and predictors. Dunn and Smyth transform the position to the normal scale [@dunnsmyth1996], whereas DHARMa obtains it by simulating from the fitted model and keeps it on the 0-to-1 scale [@hartig2024]. We use the DHARMa implementation in R.

AIC compares the models it is given. Our candidate models share the response, the family, and the structural formula and differ only in the link, so they have the same number of parameters and AIC ranks them by log-likelihood alone.

The three checks share two limits. A significant Pregibon term or a departure from uniformity can arise from a wrong family, an omitted nonlinearity, or a missing interaction as well as from a wrong link, and AIC can only rank the links that were fitted. In addition, every fitted model already contains the product term, which can absorb part of the curvature that a wrong link produces, so a misspecified model can fit almost as well as the correct one while placing additivity on a different scale.

## Diagnostic Simulation

The five scenarios are based on examples introduced earlier in the paper, with sample sizes and coefficients reported in Supplement A. Each keeps the outcome family and the interaction formula fixed and changes only the link, so the fitted model is wrong about the scale on which effects add and about nothing else. *Count errors* follows the Poisson-log example of @fig-simple-count-example and fits it with an identity link. *Chance-floor accuracy* takes the lower-performance scenario of Simulation 1, including its trial-level responses and subject random intercept, and compares standard-logit and chance-corrected-logit GLMMs with the same random-effects structure. *Binary repeated trials* follows the probit-generated, logit-fitted case of Simulation 2 with `r dg_N_binary` participants instead of `r wf_N`. *Binary continuous predictor* repeats that case with a continuous predictor in place of the two-level condition, so that residual structure can be examined along a predictor. *Gamma mean response time* generates response times from a Gamma model with a log link and fits the inverse link.

In each replication we applied the checks to the misspecified interaction model, since this is the model an analyst would inspect after obtaining a significant interaction. The residual checks covered uniformity, dispersion, quantile structure over fitted values, and structure over the focal predictor, the last assessed with a quantile check for a continuous predictor and with a comparison across design cells in the $2 \times 2$ case. AIC compared the misspecified model with the target model. The residual checks and the Pregibon test were also applied to the correctly specified model fitted to the same data, to establish how often each flags a model that is right. All rates are computed over the replications in which the misspecified model converged, which was all `r n_rep_diagnostic` replications except in the count scenario, where the Poisson-identity model converged in `r dg_n("Count errors")`.

```{r}
#| label: tbl-diagnostic-scenarios
#| tbl-cap: "Diagnostic simulation across five link-misspecification scenarios. Pseudo-interaction reports how often the misspecified model detected an interaction that was absent on the generating link scale. DHARMa gives the range across the four prespecified residual checks. Pregibon reports the added-term link check. AIC favors target reports the proportion of replications in which the lower AIC favored the target interaction model over the misspecified interaction model. CI = Wilson 95% confidence interval. The Pregibon check is not identifiable in the saturated $2 \\times 2$ design, where the squared linear predictor is exactly collinear with the existing fixed effects."
#| echo: false

knitr::kable(
  diagnostic_scenario_display_table(diagnostic_scenarios),
  align = c("l", "l", "l", "r", "r", "r", "r")
)
```

The diagnostic tables must be regenerated after converting the chance-floor scenario to the revised subject-random-intercept design. Until then, the saved D2 rates from the earlier fixed-effect design are not results for the model described here. The other diagnostic scenarios are unchanged.

# From Link Choice to Sensitivity Analysis

The practical implication of this framework is that an interaction analysis should make explicit the three choices it rests on, the outcome family, the link that defines additivity, and the metric the claim refers to, and should test whether its conclusion survives plausible alternatives (@tbl-link-aware-workflow). The observed metric may sometimes be the relevant target. For example, a two-point difference on a clinical or educational score may be the quantity that matters in practice, and additivity in raw-score units may then be reasonable. What matters is that this is a substantive choice, made deliberately and reported. When possible, the link should be justified from theory, task structure, or measurement.

When link choice can affect the interaction conclusion, we suggest that the main robustness check is a sensitivity analysis across plausible links. Researchers should therefore compare the same target response-scale contrast across alternatives and report whether the result is stable or link-conditional. If it changes, that dependence is itself part of the result [@mccabe2022; @mize2019; @rohrer2026; @mustillo2018].

For factorial designs with cell-level data, the monotonic linear model of @dunnanderson2026 offers a more general test of scale dependence, asking whether additivity can be recovered under any monotonic transformation, and every pseudo-interaction simulated here is removable in that sense. Removability alone does not decide which transformations are scientifically admissible, because the family, the bounds, the chance level, or the substantive interpretation of an outcome usually restrict the plausible mappings. Our question is narrower, namely whether interaction conclusions change across the standard link functions that remain plausible for a given outcome. 
 
```{r}
#| label: tbl-link-aware-workflow
#| tbl-cap: "Practical starting points for link-aware interaction testing across common psychological outcomes."
#| tbl-colwidths: [20, 42, 38]
#| echo: false

workflow <- data.frame(
  `Outcome type` = c(
    "Forced-choice accuracy",
    "Binary / proportion outcome",
    "Likert / ordinal item",
    "Sum score / composite",
    "Counts / error counts",
    "Response time / skewed positive outcome"
  ),
  `Plausible starting point` = c(
    "Binomial logit/probit; chance-corrected link when the task floor is part of the estimand.",
    "Binomial logit/probit; consider cloglog for asymmetric processes.",
    "Ordinal logit/probit; metric model only with a clear observed-scale target.",
    "Gaussian identity when many levels stay far from bounds; otherwise consider item-level, ordinal, bounded-count, or bounded-score models.",
    "Poisson or negative binomial, usually with log link; identity only with an observed-count target.",
    "Lognormal, Gamma log/inverse, Inverse Gaussian, or transformed Gaussian, matched to absolute, proportional, reciprocal, or transformed change."
  ),
  `Minimal sensitivity check` = c(
    "Compare standard logit/probit and chance-corrected links; report convergence and response-scale contrasts.",
    "Compare logit/probit; inspect calibration and contrasts near bounds or rare-event regions.",
    "Compare ordinal and metric summaries; report category probabilities or response-scale contrasts.",
    "Inspect bounds, skew, and score concentration; compare with item-level or bounded-score sensitivity models when feasible.",
    "Check overdispersion and zero inflation; compare count contrasts across plausible count models.",
    "Compare raw-, log-, and response-scale summaries; inspect tail fit and predicted values."
  ),
  check.names = FALSE
)

knitr::kable(workflow, align = c("l", "l", "l"))
```

# Discussion

The same data can support different interaction conclusions under different plausible specifications, which is the familiar scale dependence of interactions [@cox1984; @dunnanderson2026; @loftus1978; @wagenmakers2012]. In a GLM that dependence is carried by two choices, the outcome family and the link function. Our review found interaction tests to be common in all five journals sampled, the link function usually left implicit, and identity-link analyses applied to many outcomes for which observed-scale additivity is not self-evident.

The simulations then showed what that can cost. With two main effects present, a link that misplaces additivity turned a purely additive generating process into a significant product term even when the outcome family was correct. In forced-choice accuracy the binomial family is correct and the error lies entirely in a canonical link that places the lower asymptote at 0 rather than at the chance level. The logit-probit comparison showed that the problem survives when the response bounds are correct as well, between two links that look alike over most of the range. In the sum-score case the choice extends beyond the link to the outcome itself. A total is a scoring rule applied to item responses, so an interaction estimated on it refers to raw-score points, and adopting a family that respects the bounds and the discreteness of the total still leaves open whether additivity in points stands for additivity in the construct.

This difficulty is of a different kind from the one documented in work on the precision of interaction estimates [@baranger2023; @castillo2026]. There the obstacle is noise, and more data helps. Here the product term induced by the wrong link is a fixed property of that scale, so a larger sample estimates it more precisely and reports it as significant more often.

The diagnostic simulation asks whether ordinary post-fit checks catch link misspecification. Residual checks, Pregibon-style tests, and AIC answer different questions, and AIC can only rank the links it is given. Fresh diagnostic results are required for the revised chance-floor mixed-model scenario. More generally, these diagnostics answer a question about fit within a supplied candidate set, which is different from whether an interaction claim holds on the scale the researcher means [@dunnsmyth1996; @hartig2024; @pregibon1980].

Three limitations bound the scope of these conclusions. The review was descriptive and covered five prominent journals across major areas of psychology in a single year, so its estimates describe reporting practice in such outlets and leave the field as a whole unmeasured, and a broader mapping across journals, years, and subfields would be needed to support reporting guidance. The review also coded observable reporting features and did not assess removability in each article. The simulations, finally, are demonstrations of a mechanism. They target a single regime, in which two established main effects push expected values into the region where link curvature is strongest, and they say nothing about how often that regime arises or about the further complications of real data.

To conclude, interaction claims rest on a scale that is rarely stated. Naming it separates three questions that a software default otherwise settles at once, which family suits the response, which link defines additivity, and which metric the claim is about. The robustness question then takes a definite form, whether the conclusion survives the links that remain defensible. A result that holds across them is robust to that choice. One that does not is still a result, and it should be reported as conditional on the scale on which it was estimated.

### Author Contributions
Author roles were classified using the Contributor Roles Taxonomy (CRediT; <https://credit.niso.org/>) as follows — Conceptualization: E. Toffalini, F. Gambarota, T. Feraco; Investigation: M. Calderan, L. Sità, E. Toffalini, F. Gambarota, T. Feraco; Formal Analysis: L. Sità, E. Toffalini, F. Gambarota; Methodology: E. Toffalini, F. Gambarota; Software: E. Toffalini, F. Gambarota; Writing - Original Draft: M. Calderan, E. Toffalini; Writing - Review & Editing: F. Gambarota, T. Feraco, and L. Sità; Visualization: E. Toffalini. All authors approved the final version of the manuscript and agree to be accountable for the work.

### Conflicts of Interest
The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.

### Acknowledgments
AI tools were used for both code optimization and manuscript editing.

### Data, Code and Materials
All data, code, and supplementary files are accessible at <https://zeonod......>.

### Prior Versions
A version of this manuscript is posted as a preprint on PsyArXiv at <https://.....>. The submitted manuscript shares its full content with that preprint.

# References

::: {#refs}
:::
