---
title: "Mostly Interactions are (probably) Wrong"
shorttitle: "Short Title in Running Header"
author:
  - name: Enrico Toffalini
    corresponding: true
    orcid: 0000-0002-1404-5133
    email: enrico.toffalini@unipd.it
    affiliations:
      - name: University of Padova
        department: Department of General Psychology
        address: Via Venezia 8
        city: Padova
        country: Italy
        postal-code: 35131
        id: dpg
  - name: Filippo Gambarota
    orcid: 0000-0002-6666-1747
    email: filippo.gambarota@unipd.it
    affiliations:
      - name: University of Padova
        department: Department of Developmental Psychology and Socialization
        address: Via Venezia 8
        city: Padova
        country: Italy
        postal-code: 35131
        id: dpss
  - name: Laura Sità
    orcid: 0009-0000-6017-9979
    email: laura.sita@unipd.it
    affiliations:
      - ref: dpg
  - name: Tommaso Feraco
    orcid: 0000-0002-8920-5330
    email: tommaso.feraco@unipd.it
    affiliations:
      - ref: dpg
  - name: Margherita Calderan
    orcid: 0009-0005-5668-5162
    email: margherita.calderan@unipd.it
    affiliations:
      - ref: dpg
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "This document is a template."
#impact-statement: "This template helps you get started."
keywords: [Keyword 1, Keyword 2, Keyword 3]
bibliography: references.bib
format:
  # apaquarto-docx: default
  # apaquarto-html: default
  # apaquarto-typst: default
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  dev = "pdf",
  fig.align = "center"
)
```

```{r}
#| label: packages
#| include: false

library(ggplot2)
```

```{r}
#| label: ggplot-setup

mtheme <- function(size = 15){
  ggplot2::theme_bw(size)
}

theme_set(mtheme())
```


<!-- The introduction should not have a level-1 heading such as Introduction. -->

This is my first paragraph. Any section headings in the introduction should be in levels 2--5.

# Introduction

@Liddell2018-wu and @Wagenmakers2012-tk provide a very clear introduction about the problem of testing interactions with the wrong model.

# Method

# Results

The @fig-example summarise the results of the paper.

```{r}
#| label: fig-example
#| fig-cap: This is a really amazing figure.

n <- 100
x <- rnorm(n)
y <- x*0.5 + rnorm(n)

data.frame(x, y) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_smooth()
```

# Discussion

## Limitations and Future Directions

## Conclusion

# References

::: {#refs}
:::