---
title: "The Link Function Problem in Psychological Interaction Testing"
shorttitle: "Link Functions and Interactions"
author:
  - name: Margherita Calderan
    orcid: 0009-0005-5668-5162
    email: margherita.calderan@unipd.it
    affiliations:
      - name: University of Padova
        department: Department of General Psychology
        address: Via Venezia 8
        city: Padova
        country: Italy
        postal-code: 35131
        id: dpg
  - name: Filippo Gambarota
    orcid: 0000-0002-6666-1747
    email: filippo.gambarota@unipd.it
    affiliations:
      - name: University of Padova
        department: Department of Developmental Psychology and Socialization
        address: Via Venezia 8
        city: Padova
        country: Italy
        postal-code: 35131
        id: dpss
  - name: Laura Sità
    orcid: 0009-0000-6017-9979
    email: laura.sita@unipd.it
    affiliations:
      - ref: dpg
  - name: Tommaso Feraco
    orcid: 0000-0002-8920-5330
    email: tommaso.feraco@unipd.it
    affiliations:
      - ref: dpg
  - name: Enrico Toffalini
    corresponding: true
    orcid: 0000-0002-1404-5133
    email: enrico.toffalini@unipd.it
    affiliations:
      - ref: dpg
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "Equal differences on one scale become different differences on another scale. This is crucial for interaction testing because an interaction is a difference between differences, and in psychological research the observed response and the construct of interest are often expressed on different scales. Therefore, the same data can show an interaction on one scale and not on another. Generalized Linear Models make this problem explicit by separating two decisions: the family, which describes the nature of the response variable, and the link function, which sets the scale on which effects are additive. Because a model without an interaction term is additive on the link scale, choosing a link is choosing what 'no interaction' means. The paper has two connected parts. First, a preregistered review of recent articles from five prominent psychology journals shows that interaction testing is routinely performed in research, that the link function is most often unspecified, and that constrained outcomes are frequently analyzed with gaussian identity-link models. Second, worked examples and simulations develop three common response formats: forced-choice accuracy with a chance floor, within-family link choices (logit vs probit), and sum scores as a measurement case, and quantify how a plausible-looking link can turn an additive structure on one scale into a statistically significant interaction on another. A final diagnostic simulation shows that residual checks, a link test, and AIC comparisons give uneven protection and do not by themselves decide which scale the scientific claim concerns. Interaction claims remain underspecified until researchers state, justify, and probe the scale on which no interaction is assumed."
#impact-statement: "This template helps you get started."
keywords: [interactions, moderation, generalized linear models, link functions, model misspecification]
bibliography: references.bib
floatsintext: true
format:
  # apaquarto-docx: default
  # apaquarto-html: default
  # apaquarto-typst: default
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    fig-pos: "h"
    header-includes:
      - \usepackage{float}
      - \raggedbottom
    include-in-header:
      text: |
        % Single-space table bodies so tables fit in text (floatsintext).
        % Wrapped lines inside a cell are set slightly tighter than normal and
        % each row gets a little extra vertical air, so a line break inside a
        % cell is not mistaken for the start of a new table row.
        \usepackage{etoolbox}
        \usepackage{array}
        \AtBeginEnvironment{longtable}{%
          \linespread{0.95}\selectfont
          \setlength{\extrarowheight}{6pt}}
        \AtBeginEnvironment{tabular}{%
          \linespread{0.95}\selectfont
          \setlength{\extrarowheight}{6pt}}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  dev = "pdf",
  fig.align = "center"
)
```

```{r}
#| label: packages
#| include: false

library(ggplot2)
```

```{r}
#| label: ggplot-setup

plot_utils_candidates <- c(
  "../R/utils-plots.R",
  "R/utils-plots.R"
)

plot_utils_path <- plot_utils_candidates[file.exists(plot_utils_candidates)][1]

if (!is.na(plot_utils_path)) {
  source(plot_utils_path)
}

if (!exists("link_theme")) {
  link_palette <- function(keys = NULL) {
    pal <- c(
      "Standard logit" = "#0072B2",
      "Chance-corrected logit" = "#D55E00",
      "Group 0" = "#0072B2",
      "Group 1" = "#D55E00"
    )
    if (is.null(keys)) return(pal)
    out <- pal[keys]
    out[is.na(out)] <- "grey30"
    out
  }
  
  link_linetypes <- function(keys = NULL) {
    ltys <- c(
      "Standard logit" = "solid",
      "Chance-corrected logit" = "longdash",
      "Group 0" = "solid",
      "Group 1" = "longdash"
    )
    if (is.null(keys)) return(ltys)
    out <- ltys[keys]
    out[is.na(out)] <- "solid"
    out
  }
  
  link_scale_color_discrete <- function(..., values = link_palette()) {
    ggplot2::scale_color_manual(values = values, ...)
  }
  
  link_scale_linetype_discrete <- function(..., values = link_linetypes()) {
    ggplot2::scale_linetype_manual(values = values, ...)
  }
  
  percent_labels <- function(accuracy = 1) {
    force(accuracy)
    function(x) paste0(round(100 * x / accuracy) * accuracy, "%")
  }
  
  link_theme <- function(base_size = 10, base_family = "") {
    ggplot2::theme_minimal(base_size = base_size, base_family = base_family) +
      ggplot2::theme(
        plot.title = ggplot2::element_text(face = "bold", size = base_size + 1),
        plot.subtitle = ggplot2::element_text(size = base_size - 1, color = "grey25"),
        axis.title = ggplot2::element_text(size = base_size),
        axis.text = ggplot2::element_text(size = base_size - 1, color = "grey20"),
        strip.text = ggplot2::element_text(face = "bold", size = base_size - 1),
        legend.position = "bottom",
        legend.title = ggplot2::element_text(size = base_size - 1),
        legend.text = ggplot2::element_text(size = base_size - 1),
        panel.grid.minor = ggplot2::element_blank(),
        panel.grid.major = ggplot2::element_line(linewidth = 0.25, color = "grey88"),
        plot.margin = ggplot2::margin(6, 8, 6, 8)
      )
  }
}

ggplot2::theme_set(link_theme())
```

##

We start with a very simplified example. Two populations of children (e.g., clinical vs neurotypical) differ in their overall error rates on a task, and errors decrease with age in both. @fig-simple-count-example samples simulated data from such a scenario. In the left panel (observed response), errors decline with age in both groups, and the difference between groups is much larger at age six than at age ten. Is there an age-by-group interaction here?

```{r}
#| label: fig-simple-count-example
#| fig-cap: "Simulated error counts for two groups of children aged 6 to 10, shown on the observed count scale (left) and on the log scale (right). The two groups have exactly the same age effect on the log of the expected error count, so on the log scale the two fitted lines are parallel, with a constant group gap at every age. Mapped back onto the count scale, the same model produces converging curves, so the group difference shrinks with age. Whether this pattern is an interaction depends on the scale on which effects are assumed to add: on the count scale the group difference changes with age; on the log scale it does not."
#| fig.width: 7
#| fig.height: 3.2

set.seed(3)

N <- 250
d_count <- data.frame(
  age = round(runif(N, 6, 10), 1),
  condition = factor(rbinom(N, 1, .5), labels = c("Group A", "Group B"))
)

d_count$eta <- with(d_count, 2.6 - 0.55 * (age - 6) + 0.65 * (condition == "Group B"))
d_count$errors <- rpois(N, exp(d_count$eta))

fit_count <- glm(errors ~ age + condition, family = poisson(link = "log"), data = d_count)

count_pred <- expand.grid(age = seq(6, 10, length.out = 200), condition = levels(d_count$condition))
p <- predict(fit_count, newdata = count_pred, type = "link", se.fit = TRUE)

scale_levels <- c("Count scale: errors", "Log scale: log(errors)")

pred_two <- rbind(
  data.frame(
    count_pred, scale = scale_levels[1],
    fit = exp(p$fit),
    lower = exp(p$fit - 1.96 * p$se.fit),
    upper = exp(p$fit + 1.96 * p$se.fit)
  ),
  data.frame(
    count_pred, scale = scale_levels[2],
    fit = p$fit,
    lower = p$fit - 1.96 * p$se.fit,
    upper = p$fit + 1.96 * p$se.fit
  )
)
pred_two$scale <- factor(pred_two$scale, levels = scale_levels)

d_pos <- d_count[d_count$errors > 0, ]
points_two <- rbind(
  data.frame(age = d_count$age, condition = d_count$condition, y = d_count$errors, scale = scale_levels[1]),
  data.frame(age = d_pos$age, condition = d_pos$condition, y = log(d_pos$errors), scale = scale_levels[2])
)
points_two$scale <- factor(points_two$scale, levels = scale_levels)

ggplot(points_two, aes(age, y, color = condition, shape = condition)) +
  geom_point(size = 2.2, alpha = .5) +
  geom_ribbon(
    data = pred_two,
    aes(x = age, ymin = lower, ymax = upper, fill = condition, group = condition),
    inherit.aes = FALSE, alpha = .18, color = NA
  ) +
  geom_line(
    data = pred_two,
    aes(x = age, y = fit, linetype = condition, group = condition),
    linewidth = 1.1
  ) +
  facet_wrap(~ scale, scales = "free_y") +
  scale_color_manual(values = c("Group A" = "#0072B2", "Group B" = "#D55E00")) +
  scale_fill_manual(values = c("Group A" = "#0072B2", "Group B" = "#D55E00")) +
  labs(x = "Age", y = NULL, color = NULL, fill = NULL, shape = NULL, linetype = NULL) +
  link_theme(base_size = 11) +
  theme(legend.position = "top")
```

Because an interaction is a difference between differences, the question is whether the group difference itself changes with age. On the observed count scale, the group effect does appear to shrink as age increases, so it looks like an interaction. But looking at the same data on the log-scale (@fig-simple-count-example, right panel), both groups have the same age slope, meaning their fitted lines are parallel. A constant difference on the log scale corresponds to a constant ratio on the count scale, so the absolute gap between groups is proportional to the baseline and shrinks as errors decline with age. Whether we call this an interaction is not a fact about the data, it depends on which scale we assume effects add up linearly on. Additive on the log scale means multiplicative (and non-parallel) on the count scale, and vice versa.

This ambiguity extends far beyond this single illustration, as interaction hypotheses are widespread in psychological research and often represent the central target of inquiry [@rohrer2021; @domingue2024; @mccabe2022]. Whenever researchers ask whether an effect is moderated, or whether it holds only under certain  conditions, they are asserting that the impact of one variable depends on another. This count example therefore demonstrates that the reality of such a claim can shift entirely based on the scale on which those differences are computed.

In psychology, this scale dependence has a characteristic source in measurement. Researchers routinely conceptualize constructs such as cognitive ability or symptom severity as continuous latent dimensions [@embretsonReise2000; @wagenmakers2012]. Those dimensions are rarely observed directly; what reaches the data are bounded, discrete quantities, such as counts, ratings, and accuracy scores that carry floors, ceilings, and chance levels. Each measurement scale imposes its own mapping from the latent dimension to the observed response. Where the mapping compresses, a constant effect on the latent dimension appears as a shrinking effect on the observed response: a group or condition closer to a bound has less room to move, and can look less affected by a manipulation, than a group or condition farther away from bounds (e.g., @fig-simple-count-example, left panel). 

Generalized linear models (GLMs) make this mapping explicit and give it a standard vocabulary [@nelderwedderburn1972; @mccullaghnelder1989], separating two choices that are easy to conflate, (1) the outcome family, which describes the response, and (2) the link function, which sets the scale on which effects combine (Box 1).

::: {.callout-note title="Box 1: Generalized linear models: outcome family and link function" icon="false"}
The *outcome family* is the distribution assumed for the response. It defines which values the outcome can take and how observations vary around their expectation. The *link function*, $g(\cdot)$, maps the expected outcome, $\mu_i = E(Y_i)$, onto the linear predictor, $\eta_i$, the scale on which the model is additive. With two predictors and their interaction term, $$g(\mu_i) = \eta_i, \qquad \eta_i = \beta_0 + \beta_1 X_i + \beta_2 Z_i + \beta_3 X_i Z_i.$$

The inverse link, $g^{-1}(\cdot)$, maps this additive structure back onto the expected response. Its familiar role is to keep expected values within the admissible range of a constrained outcome: $\exp(\eta)$ keeps expected counts positive, and the inverse logit keeps probabilities between 0 and 1. Less obviously, this also fixes what the absence of an interaction, $\beta_3 = 0$, means in the fitted model. The product-term coefficient $\beta_3$ is therefore an interaction on the link scale. A response-scale interaction is a different quantity: a difference between expected differences on the observed outcome, for example, whether the group difference in expected errors changes with age in @fig-simple-count-example. Unless the link is the identity, the two quantities can disagree [@ainorton2003; @mccabe2022; @mize2019]. Each family has a canonical link, which is the one that arises directly from the family's mathematical form, but it is a convenient default and other links are legitimate. A Gaussian linear model is the special case that combines a Gaussian family with an identity link, so the additive scale and the response scale coincide and the scale question is invisible. The model behind @fig-simple-count-example is a Poisson family with a log link. The family respects that errors are non-negative counts, and the link places additivity on the log scale.
:::

Often the outcome family is chosen with care while the link function is left at its software default, fixing the scale of additivity without deliberation. Two models can share a family and still place additivity on different scales. For example, a Gamma family suits reaction times data; within that family, a log link makes effects additive in log milliseconds, a proportional change, and the canonical inverse link makes them additive in reciprocal means, a change in speed. A binomial family fits correct-versus-incorrect responses and leaves the link open among logit, probit, chance-corrected, and other forms.

The scale dependence of interactions has long been recognized. Loftus showed that interactions involving response probabilities are meaningful only relative to a mapping between observed and theoretical scales [@loftus1978], and Cox formalized interaction as a model-relative concept tied to transformation and interpretation [@cox1984]. Busemeyer and Jones carried the same point into the measurement model, where the conclusion about how two variables combine depends on the model assumed to link responses to the quantities they measure [@busemeyer1983]. Lubinski and Humphreys showed that the specification choices alone can produce the appearance of an interaction where the generating process contains none [@lubinski1990].

More recently, Wagenmakers and colleagues separated removable from nonremovable interactions [@wagenmakers2012]. A removable interaction can be undone by a monotonic transformation of the dependent variable, which makes it a property of the measurement scale, whereas a nonremovable one (typically a cross-over) persists under any such transformation. Rohrer and Arslan showed that the same interaction can be positive, null, or negative depending on the scale chosen for the outcome, so the appropriate test follows from the scale on which the hypothesis is posed [@rohrer2021]. Rimpler and colleagues added that a product term assumes one predictor's effect changes at a constant rate across levels of the other, and so misrepresents any dependence that does not take that form, such as a threshold effect [@rimpler2026].  In GLMs specifically, the product-term coefficient refers to the link scale and does not equal the interaction on the response scale, so a question about the response scale should be answered with marginal effects [@ainorton2003; @mccabe2022; @mize2019; @rohrer2026]. 

Domingue and colleagues demonstrated that outcome-model misspecification can produce false interaction findings for binary, count, and censored outcomes [@domingue2024], and Liddell and Kruschke showed that treating ordinal responses as metric can inflate error rates for interactions [@liddellkruschke2018]. Czado and Santner, and later Yu and Huang, studied the inferential consequences of link misspecification in binary GLMs and GLMMs [@czadosantner1992; @yuhuang2019]. Beyond the scale question, interaction estimates are fragile for other reasons, including small effects, low reliability, and limited power [@baranger2023; @castillo2026]. 

This literature spans nearly five decades, but whether applied practice has absorbed the warning is still unaddressed. The present article approaches that gap from two directions. We first document current practice with a preregistered review of recent empirical articles in five prominent psychology journals, quantifying how often interactions are tested, how often the outcome family and link function are made explicit, and which response formats predominate. We then develop a framework that separates outcome family, link function, and target metric, and work it through three common response formats, namely forced-choice accuracy with a chance floor, sum scores, and within-family link choices such as logit versus probit, using simulations to quantify how often an inadequate link turns a purely additive structure into a statistically significant interaction and whether standard diagnostics detect the misspecification.

These simulations take up the case in which the outcome family is already broadly plausible and the link alone sets a disputable no-interaction baseline, the common situation for bounded and chance-limited psychological outcomes. Work on misspecification has largely varied the outcome model as a whole [@domingue2024; @liddellkruschke2018], whereas here the family stays fixed and the link carries the entire disagreement about what counts as no interaction. An interaction that appears on a scale other than the generating one is then a genuine mathematical property of that scale [@loftus1978; @cox1984; @domingue2024; @mccabe2022], correct where it was computed and troublesome only when the claim is carried to a different scale. Because the generating scale is known in a simulation, we can call such a result a pseudo-interaction, an interaction that is zero on the generating scale and nonzero on another fitted scale because the two encode different additivity baselines. In empirical data, where the generating scale is unknown, we describe interaction conclusions as link-sensitive, conditional on the fitted scale, or unstable across plausible links. @tbl-link-aware-workflow summarizes these insights into a practical workflow for applied research.

Technical details are reported in Supplement A; a broader precomputed sensitivity atlas varying sample size, main-effect magnitude, the intercept, and selected design parameters is reported in Supplement B. 

# Current Practice in Psychological Interaction Testing: A Preregistered Review

```{r review-section-load-data}
#| include: false

find_repo_file <- function(...) {
  rel <- file.path(...)
  candidates <- c(rel, file.path("..", rel))
  hit <- candidates[file.exists(candidates)]
  if (length(hit) == 0) {
    stop(
      "Cannot find required file: ", rel, ".\n",
      "Looked in: ", paste(candidates, collapse = "; "), ".\n",
      "Run the corresponding script before rendering paper.qmd."
    )
  }
  hit[1]
}

review_table_path <- find_repo_file("tables", "review-summary.csv")
outcome_table_path <- find_repo_file("tables", "review-outcome-types.csv")
agreement_table_path <- find_repo_file("tables", "review-intercoder-agreement.csv")
journal_table_path <- find_repo_file("tables", "review-by-journal.csv")
screening_flow_path <- find_repo_file("tables", "review-screening-flow.csv")

review_table <- read.csv(review_table_path)
outcome_table <- read.csv(outcome_table_path)
agreement_table <- read.csv(agreement_table_path)
journal_table <- read.csv(journal_table_path)
screening_flow_table <- read.csv(screening_flow_path)

get_review_row <- function(row_id, data = review_table) {
  if (!"row_id" %in% names(data)) {
    stop("Review summary table must contain a row_id column. Re-run scripts/01-review-descriptives.R before rendering.")
  }
  idx <- which(data$row_id == row_id)
  if (length(idx) != 1) {
    stop("Could not uniquely identify review row_id: ", row_id)
  }
  data[idx, ]
}

n_eligible <- get_review_row("eligible_empirical")
n_interactions <- get_review_row("testing_interactions")
n_non_identity <- get_review_row("non_identity_link")
n_explicit <- get_review_row("explicit_link")
n_incorrect_identity <- get_review_row("incorrect_identity")
n_significant_incorrect_identity <- get_review_row("significant_incorrect_identity")

get_outcome_row <- function(pattern, data = outcome_table) {
  idx <- grep(pattern, data$outcome_type, ignore.case = TRUE)
  if (length(idx) != 1) {
    stop("Could not uniquely identify outcome row: ", pattern)
  }
  data[idx, ]
}

get_agreement_row <- function(variable, data = agreement_table) {
  idx <- which(data$variable == variable)
  if (length(idx) != 1) {
    stop("Could not uniquely identify agreement variable: ", variable)
  }
  data[idx, ]
}

get_journal_row <- function(journal, data = journal_table) {
  idx <- which(data$journal == journal)
  if (length(idx) != 1) {
    stop("Could not uniquely identify journal row: ", journal)
  }
  data[idx, ]
}

outcome_binary <- get_outcome_row("Binary")
outcome_sum <- get_outcome_row("Sum scores")
outcome_ordinal <- get_outcome_row("Ordinal")
outcome_rt <- get_outcome_row("Response times")
outcome_neural <- get_outcome_row("Neural")
outcome_difference <- get_outcome_row("Difference")
outcome_correlations <- get_outcome_row("Correlations")
outcome_counts <- get_outcome_row("Counts")

agree_non_identity <- get_agreement_row("Uses_non_identity_link_function")
agree_explicit <- get_agreement_row("Explicit_link_function")
agree_incorrect_identity <- get_agreement_row("Incorrect_identity_link_function")
agree_significant <- get_agreement_row("Finds_significant_interaction")

journal_psychmed <- get_journal_row("Psychological Medicine")
journal_min_incorrect_identity <- journal_table[which.min(journal_table$incorrect_identity_percent_interaction), ]
journal_max_incorrect_identity <- journal_table[which.max(journal_table$incorrect_identity_percent_interaction), ]
journal_min_interaction <- journal_table[which.min(journal_table$interaction_testing_percent_of_eligible), ]
journal_max_interaction <- journal_table[which.max(journal_table$interaction_testing_percent_of_eligible), ]

sem_row_idx <- grep("SEM|latent-variable-only", screening_flow_table$step, ignore.case = TRUE)
if (length(sem_row_idx) != 1) {
  stop("Could not uniquely identify SEM / latent-variable-only exclusions row in screening_flow_table")
}
sem_exclusions_n <- screening_flow_table$n[[sem_row_idx]]

fmt_ci <- function(row) {
  paste0(row$percent, "%, 95% CI [", row$ci_low, ", ", row$ci_high, "]")
}

fmt_num <- function(x, digits = 3) {
  out <- ifelse(is.na(x), "--", formatC(as.numeric(x), format = "f", digits = digits))
  out
}

fmt_count <- function(x) {
  ifelse(is.na(x), "--", formatC(as.numeric(x), format = "f", digits = 0))
}

fmt_agreement_kappa <- function(row) {
  paste0(
    fmt_num(100 * row$agreement, 1), "% agreement, kappa = ",
    fmt_num(row$kappa, 2)
  )
}

fmt_yes_margins <- function(row) {
  if (all(c("coder1_yes_percent", "coder2_yes_percent") %in% names(row))) {
    return(paste0(
      fmt_num(row$coder1_yes_percent, 1), "% / ",
      fmt_num(row$coder2_yes_percent, 1), "%"
    ))
  }
  n <- row$n_double_coded
  coder1_yes <- 100 * (row$n_10 + row$n_11) / n
  coder2_yes <- 100 * (row$n_01 + row$n_11) / n
  paste0(fmt_num(coder1_yes, 1), "% / ", fmt_num(coder2_yes, 1), "%")
}

fmt_ci_cols <- function(low, high, digits = 3) {
  ifelse(
    is.na(low) | is.na(high),
    "--",
    paste0("[", fmt_num(low, digits), ", ", fmt_num(high, digits), "]")
  )
}

review_display_table <- function(dat) {
  out <- data.frame(
    Quantity = dat$quantity,
    n = fmt_count(dat$n),
    `Denom.` = fmt_count(dat$denominator_n),
    `%` = paste0(fmt_num(dat$percent, 1), "%"),
    `95% CI` = paste0("[", fmt_num(dat$ci_low, 1), "%, ", fmt_num(dat$ci_high, 1), "%]"),
    check.names = FALSE
  )
  out
}

simulation_display_table <- function(dat, effect_col = NULL, include_scale = FALSE) {
  out <- data.frame(.row = seq_len(nrow(dat)), stringsAsFactors = FALSE)
  if ("Scenario" %in% names(dat)) out$Scenario <- dat$Scenario
  if ("Generating link" %in% names(dat)) out$`Generating link` <- dat$`Generating link`
  if ("Model" %in% names(dat)) out$Model <- dat$Model
  if ("Fitted link" %in% names(dat)) out$`Fitted link` <- dat$`Fitted link`
  if ("Link status" %in% names(dat)) out$`Link status` <- dat$`Link status`
  if ("Rate type" %in% names(dat)) out$`Rate type` <- dat$`Rate type`
  if (include_scale && "Units for group-gap change" %in% names(dat)) {
    out$Scale <- ifelse(grepl("latent", dat$`Units for group-gap change`), "latent theta", "sum score")
  }
  if ("Successful fits" %in% names(dat)) out$Fits <- fmt_count(dat$`Successful fits`)
  if ("Rejections" %in% names(dat)) out$Reject <- fmt_count(dat$Rejections)
  if ("Rejection rate" %in% names(dat)) {
    out$`Rejection rate` <- fmt_num(dat$`Rejection rate`, 3)
  }
  if (all(c("CI low", "CI high") %in% names(dat))) {
    out$`95% CI` <- fmt_ci_cols(dat$`CI low`, dat$`CI high`, 3)
  }
  if (!is.null(effect_col) && effect_col %in% names(dat)) {
    out$`Median gap change` <- fmt_num(dat[[effect_col]], 3)
  }
  if ("Median interaction coefficient" %in% names(dat)) {
    out$`Median beta interaction` <- fmt_num(dat$`Median interaction coefficient`, 3)
  }
  if ("Median interaction SE" %in% names(dat)) {
    out$`Median SE` <- fmt_num(dat$`Median interaction SE`, 3)
  }
  out$.row <- NULL
  out
}

rate_display_table <- function(dat) {
  data.frame(
    Quantity = dat$Quantity,
    Fits = fmt_count(dat$`Successful fits`),
    Sig = fmt_count(dat$`Significant tests`),
    Rate = fmt_num(dat$Rate, 3),
    `95% CI` = fmt_ci_cols(dat$`CI low`, dat$`CI high`, 3),
    check.names = FALSE
  )
}

```

The review asks how often psychological articles test interactions, and how often they make the outcome family and link function explicit enough for a reader to know which interaction claim was tested. To our knowledge, it is the first review to quantify link-function reporting in psychological interaction tests. The coding is descriptive, so it documents reporting practice rather than the correctness of any substantive conclusion.

## Review Method

The protocol was preregistered before any record was screened (<https://osf.io/bdu73/>), together with the coding data dictionary. The essentials are given here; the full protocol, sampling frame, and reliability are in Supplement A.

Following the preregistration, we sampled one journal from each of five areas: *Psychological Science* (general), *Journal of Experimental Psychology: General* (experimental), *Journal of Personality and Social Psychology* (social), *Developmental Science* (developmental), and *Psychological Medicine* (clinical). The five were selected on preregistered discretionary criteria: a first-quartile 2023 Impact Factor, a broad rather than narrowly specialized scope, not primarily publishing reviews, and strong reputation in their field, as jointly judged by the coauthors. From each journal's full set of 2024 articles in Scopus, records were screened in a fixed random order until 40 eligible interaction-testing articles were reached, for a preregistered total of 200. Because screening stopped once 40 interaction-testing articles were reached, the number of records screened, and therefore the denominator of the eligibility proportion, varied across journals. Eligible articles were empirical reports of original research; reviews, meta-analyses, editorials, theoretical and qualitative-only articles, replication-only studies, and psychometric validation studies were excluded. Detailed coding was applied only to articles testing at least one interaction, and only to observed-outcome analyses: articles whose relevant tests were purely latent-variable or structural-equation models fell outside the coding scheme (`r sem_exclusions_n` such exclusions).

Each interaction-testing article was independently coded by two researchers, with disagreements resolved by discussion or by consulting the other coauthors. Generative AI tools were used only as an optional aid to locate passages within an article; all coding used in the review is human.

For each article, the coders recorded five things: whether at least one tested interaction used a non-identity link; whether the link function was explicitly reported; whether at least one interaction applied an incorrect identity-link analysis; whether at least one such interaction was statistically significant, or interpreted analogously in a Bayesian analysis; and which response-variable types were analyzed. 

By a formally incorrect identity-link analysis we mean a Gaussian linear regression with an identity link applied to an outcome with formal constraints, such as bounds, discreteness, positivity, or a known chance level. This is a conservative screening flag, applied only when the constraint was identifiable from the report and does not imply the substantive conclusion is wrong. The label is defined relative to the constrained process that generates the outcome (if the observed score were explicitly justified as the target metric, identity-link additivity would instead be a defensible modeling choice, a distinction we develop in the following sections). 

The full operational rules the coders followed, including borderline cases, are tabulated in Supplement A. Interrater agreement on the double-coded records was high (Cohen's $\kappa$ between `r fmt_num(min(agreement_table$kappa), 2)` and `r fmt_num(max(agreement_table$kappa), 2)`); per-variable agreement and coder margins are reported in Supplement A. Preregistered variable definitions, operational coding rules, variable-specific agreement, and by-journal results are also reported in Supplement A, Section S2.

## Review Findings

First, interaction testing was common. Screening yielded `r n_eligible$n` eligible empirical articles, of which  `r n_interactions$n`  tested at least one interaction,  `r fmt_num(n_interactions$percent, 1)`%. Under the stopping rule this proportion describes the records screened rather than a fixed-denominator sample, so we report it descriptively and treat the per-journal proportions as the more informative summary: these ranged from  `r fmt_num(journal_min_interaction$interaction_testing_percent_of_eligible, 1)`% in  `r journal_min_interaction$journal`  to  `r fmt_num(journal_max_interaction$interaction_testing_percent_of_eligible, 1)`% in  `r journal_max_interaction$journal`. 

Second, model-scale reporting was uncommon. Among the `r n_interactions$n` interaction-testing articles, `r n_non_identity$n` used at least one non-identity link function, `r fmt_ci(n_non_identity)`, and only `r n_explicit$n` explicitly reported the link function, `r fmt_ci(n_explicit)`. Thus, in about `r fmt_num(100 - n_explicit$percent, 1)`% of interaction-testing articles, the link function was not explicitly stated.

Even the `r n_explicit$n` explicit cases rest on a rule looser than the preregistered one, which required a declaration of the link and treated naming a family or a generalized linear model as insufficient. In coding we also accepted link-specific model names, because in practice they identify the link: "logistic regression" names the logit and "probit regression" the probit. We also accepted "Poisson regression", which names a family whose standard link is the log. The stricter preregistered rule would give a lower count. 

Third, constrained outcomes were often analyzed on the identity scale. Among the `r n_interactions$n` interaction-testing articles, `r n_incorrect_identity$n` included at least one formally incorrect identity-link analysis of a constrained observed outcome, `r fmt_ci(n_incorrect_identity)`. As noted, this is a conservative screening count, flagged only when the constraint was clear from the report. Of these `r n_incorrect_identity$n` cases, `r n_significant_incorrect_identity$n` reported at least one statistically significant interaction, `r fmt_ci(n_significant_incorrect_identity)`. This is not a comparison, because significance was coded only within the flagged analyses; that is, `r fmt_num(100 * n_significant_incorrect_identity$n / n_interactions$n, 1)`% of all interaction-testing articles combined a constrained outcome, an identity-link analysis, and a significant interaction claim. This is the same concern raised by work on outcome-model mismatch in interaction testing [@domingue2024; @liddellkruschke2018]. @tbl-current-practice summarizes these results.

```{r}
#| label: tbl-current-practice
#| tbl-cap: "Descriptive summary of the preregistered review. Formally incorrect identity-link analyses of constrained observed outcomes are cases where the reported outcome had formal constraints that make identity-link additivity formally incorrect as a generative model unless explicitly justified as an approximation or as the target observed-score metric. The interval on the eligibility row is reported for completeness only: screening stopped at 40 interaction-testing articles per journal, denominator was not fixed by design."
#| echo: false

knitr::kable(
  review_display_table(review_table),
  align = c("l", "r", "r", "r", "r")
)
```

The pattern was present across the sampled journals. In Supplement A, the proportion of formally incorrect identity-link analyses of constrained observed outcomes among interaction-testing articles ranged from `r fmt_num(journal_min_incorrect_identity$incorrect_identity_percent_interaction, 1)`% in `r journal_min_incorrect_identity$journal` to `r fmt_num(journal_max_incorrect_identity$incorrect_identity_percent_interaction, 1)`% in `r journal_max_incorrect_identity$journal`. In this sample, `r journal_psychmed$journal` had the lowest screening proportion and the highest non-identity-link proportion (`r fmt_num(journal_psychmed$non_identity_link_percent_interaction, 1)`%). This journal, which uses many outcomes for which non-identity links are standard, contributed the fewest formally incorrect identity-link cases.

The outcomes used in interaction analyses were heterogeneous. The most common broad, non-exclusive categories were binary, accuracy, or proportion outcomes (`r outcome_binary$n` articles), sum scores or composites (`r outcome_sum$n` articles), ordinal, Likert, or rating outcomes (`r outcome_ordinal$n` articles), response times or durations (`r outcome_rt$n` articles), neural or physiological measures (`r outcome_neural$n` articles), difference or distance scores (`r outcome_difference$n` articles), counts or error counts (`r outcome_counts$n` articles), and correlations or associations (`r outcome_correlations$n` articles). These broad categories combine the more detailed labels used to classify the response variables during coding. The operational definitions and decision rules used to distinguish them are reported in Supplement A.

For approximately continuous outcomes, additivity on the observed scale can be a defensible target, whereas for bounded and discrete ones it is a stronger assumption, since treating them as metric distorts estimates, error rates, and interaction conclusions [@liddellkruschke2018; @burknervuorre2019]. Additionally, bounded outcomes dominated the corpus, since the first three categories together appeared in 152 of the 200 articles (76.0%). Therefore, the worked examples focus on binomial and aggregated responses, including sums of binary or ordinal items.

```{r}
#| label: tbl-review-outcome-types
#| tbl-cap: "Broad response-variable categories in interaction-testing articles. Categories are non-exclusive, so one article can contribute to more than one row."
#| echo: false

outcome_display <- data.frame(
  `Outcome type` = outcome_table$outcome_type,
  n = fmt_count(outcome_table$n),
  `Denom.` = fmt_count(outcome_table$denominator_n),
  `%` = paste0(fmt_num(outcome_table$percent, 1), "%"),
  check.names = FALSE
)
knitr::kable(outcome_display, align = c("l", "r", "r", "r"))
```

## What the Present Review Shows, What It Does Not

As the first review to quantify link-function reporting in psychological interaction tests, it suggests three points. First, interaction testing is pervasive, in every journal sampled and in `r fmt_num(n_interactions$percent, 0)`% of the eligible empirical articles screened. Second, the scale of additivity is usually left implicit: only `r fmt_num(n_explicit$percent, 0)`% named the link function, and even that count leans on counting model names (such as "logistic regression") rather than the link function per se. Third, constrained outcomes are routinely analyzed on the identity scale: `r fmt_num(n_incorrect_identity$percent, 0)`% of interaction-testing articles applied an identity-link analysis to an outcome with formal constraints, and most of these reported a significant interaction. In leading journals, then, interaction tests are common while the scale on which no interaction is assumed is rarely stated.

Nevertheless, a descriptive review cannot establish how often published interactions are scale artifacts, because that judgment requires knowing the scale on which the data-generating process is additive. The remainder of the paper therefore turns to settings where the generating process is known by construction (i.e., simulations).

# The Link Function Problem in Three Common Cases

To demonstrate the consequences of link misspecification we conduct three simulation studies. The three simulations ask whether a purely additive data-generating process on one scale is detected as a statistically significant interaction when modeled on another scale:

1.  *Forced-Choice Accuracy*: This case holds the binomial outcome family fixed and evaluates whether the link function must explicitly encode a task-implied chance floor (e.g., a .50 guessing threshold).
2.  *Within-Family Link Choice (Logit vs. Probit)*: This case holds both the binary family and the 0-to-1 bounds fixed, testing whether choosing between these two standard link functions alters the results.
3.  *Sum Scores*: This case holds the latent generating process fixed and evaluates whether the bounded total shows an interaction that the latent scale does not.

The simulation scenarios were chosen to represent plausible psychological data patterns. Each reproduces a common interaction-testing situation: two main effects that are well established and pretty large, for example a group difference combined with a condition effect or a developmental trend, and the question of whether one effect moderates the other. The two main effects were always tuned strong enough to push expected values close to a floor, ceiling, or chance level while keeping them clearly inside the observed range. This is the region where link curvature acts most strongly.

Each simulation applies the same design. Data are generated with no product term on the generating scale, and models assuming additivity on that scale and on plausible alternative scales are fitted to the same data. Every fitted model yields a product-term rejection rate. We called *nominal rejection rate* the rejection under a matched generating-scale model, whereas rejection under a mismatched-scale model is the *pseudo-interaction detection rate*. In the next section, a final simulation examines whether standard diagnostics reliably identify the misspecified link.

```{r simulation-load-results}
#| include: false

read_required_table <- function(file, script_hint) {
  path <- find_repo_file("tables", file)
  dat <- read.csv(path, stringsAsFactors = FALSE)
  if (nrow(dat) == 0) {
    stop("The table ", path, " exists but has zero rows. Re-run ", script_hint, ".")
  }
  dat
}

select_required_columns <- function(dat, cols) {
  missing <- setdiff(cols, names(dat))
  if (length(missing) > 0) {
    stop(
      "The following required columns were not found: ",
      paste(missing, collapse = ", ")
    )
  }
  dat[, cols, drop = FALSE]
}

rename_columns <- function(dat, labels) {
  names(dat)[match(names(labels), names(dat))] <- unname(labels)
  dat
}

# Inline row lookups must hit exactly one row. A label that no longer matches the
# summary tables would otherwise return a zero-length value that knitr renders as
# an empty string, silently dropping the number from the text.
require_single <- function(value, what) {
  if (length(value) != 1) {
    stop(
      "Inline lookup ", what, " matched ", length(value),
      " values instead of 1. Check that the label matches the summary table."
    )
  }
  value
}

forced_choice_sim_raw <- read_required_table(
  "simulation-summary-forced-choice.csv",
  "scripts/03a-simulation-forced-choice.R"
)
within_family_sim_raw <- read_required_table(
  "simulation-summary-within-family-links.csv",
  "scripts/03b-simulation-within-family-links.R"
)
sum_score_sim_raw <- read_required_table(
  "simulation-summary-sum-scores.csv",
  "scripts/03c-simulation-sum-scores.R"
)
diagnostic_sim_raw <- read_required_table(
  "simulation-summary-diagnostic-worked-example.csv",
  "scripts/04-diagnostic-worked-example.R"
)

diagnostic_scenarios_raw <- read_required_table(
  "simulation-summary-diagnostic-scenarios.csv",
  "scripts/04-diagnostic-worked-example.R"
)

diagnostic_calibration_raw <- read_required_table(
  "diagnostic-calibration.csv",
  "scripts/04-diagnostic-worked-example.R"
)

diagnostic_replications_path <- find_repo_file(
  "outputs", "diagnostic-simulation-replications.csv"
)
diagnostic_replications <- read.csv(
  diagnostic_replications_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

diagnostic_dharma_cols <- c(
  "dharma_uniformity_p",
  "dharma_dispersion_p",
  "dharma_quantile_fitted_p",
  "dharma_quantile_predictor_p",
  "dharma_categorical_design_p"
)
diagnostic_required_cols <- c(
  "scenario", "interaction_p", diagnostic_dharma_cols,
  "aic_fitted_minus_true"
)
if (!all(diagnostic_required_cols %in% names(diagnostic_replications))) {
  stop(
    "The diagnostic replication file is missing required columns: ",
    paste(
      setdiff(diagnostic_required_cols, names(diagnostic_replications)),
      collapse = ", "
    )
  )
}

diagnostic_calibration_required_cols <- c(
  "scenario", "diagnostic", "model_specification", "n_attempted",
  "n_successful", "flagging_rate", "ci_low", "ci_high"
)
if (!all(diagnostic_calibration_required_cols %in% names(diagnostic_calibration_raw))) {
  stop(
    "The diagnostic calibration table is missing required columns: ",
    paste(
      setdiff(
        diagnostic_calibration_required_cols,
        names(diagnostic_calibration_raw)
      ),
      collapse = ", "
    )
  )
}
diagnostic_calibration_correct <- diagnostic_calibration_raw[
  diagnostic_calibration_raw$model_specification == "Correct link" &
    diagnostic_calibration_raw$n_successful > 0 &
    is.finite(diagnostic_calibration_raw$flagging_rate),
  ,
  drop = FALSE
]
# In the saturated 2 x 2 interaction model the squared population-level linear
# predictor is collinear with the existing fixed effects, so the
# Pregibon-style check is not applicable. The continuous-predictor GLMM remains
# estimable and is calibrated separately below.
pregibon_na_scenario <- "Binary repeated trials"
pregibon_glmm_continuous_scenario <- "Binary continuous predictor"

diagnostic_calibration_dharma <- diagnostic_calibration_correct[
  grepl("^DHARMa", diagnostic_calibration_correct$diagnostic), , drop = FALSE
]
diagnostic_calibration_pregibon <- diagnostic_calibration_correct[
  grepl("Pregibon", diagnostic_calibration_correct$diagnostic), , drop = FALSE
]
dharma_calibration_max <- 100 * max(
  diagnostic_calibration_dharma$flagging_rate,
  na.rm = TRUE
)
pregibon_calibration_single_level_max <- 100 * max(
  diagnostic_calibration_pregibon$flagging_rate[
    !(diagnostic_calibration_pregibon$scenario %in% c(
      pregibon_na_scenario,
      pregibon_glmm_continuous_scenario
    ))
  ],
  na.rm = TRUE
)
pregibon_calibration_glmm_continuous <- 100 * require_single(
  diagnostic_calibration_pregibon$flagging_rate[
    diagnostic_calibration_pregibon$scenario == pregibon_glmm_continuous_scenario
  ],
  "correct-link Pregibon calibration for the continuous-predictor GLMM"
)

diagnostic_calibration_regular <- (
  nrow(diagnostic_calibration_correct) > 0 &&
    all(diagnostic_calibration_correct$ci_low <= .05, na.rm = TRUE)
)
diagnostic_calibration_result_sentence <- if (diagnostic_calibration_regular) {
  paste(
    "Baseline flagging rates under the correct link were generally low",
    "across diagnostics and scenarios; full calibration results are reported",
    "in Supplement A"
  )
} else {
  paste(
    "Calibration rates varied across diagnostics or scenarios, so flagging",
    "rates under misspecification should be read against the baseline rates",
    "reported in Supplement A."
  )
}

# Scenario tables are produced by the revised scripts and document the
# deterministic values implied by each data-generating scenario. They are
# loaded here so that the manuscript and scripts remain synchronized, but
# they are not printed as extra main-text tables. The multi-panel figures
# are the primary scenario visualizations.
forced_choice_scenario <- read_required_table(
  "scenario-table-forced-choice.csv",
  "scripts/03a-simulation-forced-choice.R"
)
within_family_scenario <- read_required_table(
  "scenario-table-within-family-links.csv",
  "scripts/03b-simulation-within-family-links.R"
)
sum_score_scenario <- read_required_table(
  "scenario-table-sum-scores.csv",
  "scripts/03c-simulation-sum-scores.R"
)
diagnostic_scenario <- read_required_table(
  "scenario-table-diagnostic-worked-example.csv",
  "scripts/04-diagnostic-worked-example.R"
)

forced_choice_sim <- select_required_columns(
  forced_choice_sim_raw,
  c(
    "scenario",
    "model",
    "n_successful_fits",
    "n_rejections",
    "rejection_rate",
    "rate_type",
    "ci_low",
    "ci_high",
    "sd_interaction_coef",
    "median_change_in_group_difference_outcome_units"
  )
)
forced_choice_sim <- rename_columns(
  forced_choice_sim,
  c(
    scenario = "Scenario",
    model = "Model",
    n_successful_fits = "Successful fits",
    n_rejections = "Rejections",
    rejection_rate = "Rejection rate",
    rate_type = "Rate type",
    ci_low = "CI low",
    ci_high = "CI high",
    sd_interaction_coef = "Monte Carlo SD interaction coefficient",
    median_change_in_group_difference_outcome_units = "Median change in group gap, correct-response units"
  )
)

sum_score_sim <- select_required_columns(
  sum_score_sim_raw,
  c(
    "scenario",
    "model",
    "change_in_group_difference_units",
    "n_successful_fits",
    "n_rejections",
    "rejection_rate",
    "rate_type",
    "ci_low",
    "ci_high",
    "median_change_in_group_difference_outcome_units"
  )
)
sum_score_sim <- rename_columns(
  sum_score_sim,
  c(
    scenario = "Scenario",
    model = "Model",
    change_in_group_difference_units = "Units for group-gap change",
    n_successful_fits = "Successful fits",
    n_rejections = "Rejections",
    rejection_rate = "Rejection rate",
    rate_type = "Rate type",
    ci_low = "CI low",
    ci_high = "CI high",
    median_change_in_group_difference_outcome_units = "Median change in group gap"
  )
)

# The within-family script writes repeated-trial GLMM results without a
# scenario column or response-scale gap-change summaries.
within_family_sim <- select_required_columns(
  within_family_sim_raw,
  c(
    "generating_link",
    "fitted_link",
    "link_match",
    "n_successful_fits",
    "n_rejections",
    "rejection_rate",
    "rate_type",
    "ci_low",
    "ci_high",
    "median_interaction_coef",
    "median_interaction_se",
    "deterministic_product_on_fitted_link_scale"
  )
)
within_family_sim <- rename_columns(
  within_family_sim,
  c(
    generating_link = "Generating link",
    fitted_link = "Fitted link",
    link_match = "Link status",
    n_successful_fits = "Successful fits",
    n_rejections = "Rejections",
    rejection_rate = "Rejection rate",
    rate_type = "Rate type",
    ci_low = "CI low",
    ci_high = "CI high",
    median_interaction_coef = "Median interaction coefficient",
    median_interaction_se = "Median interaction SE",
    deterministic_product_on_fitted_link_scale = "Induced product term"
  )
)
diagnostic_sim <- select_required_columns(
  diagnostic_sim_raw,
  c("scenario", "quantity", "n_successful_fits", "n_significant", "rate", "ci_low", "ci_high")
)

diagnostic_sim <- rename_columns(
  diagnostic_sim,
  c(
    scenario = "Scenario",
    quantity = "Quantity",
    n_successful_fits = "Successful fits",
    n_significant = "Significant tests",
    rate = "Rate",
    ci_low = "CI low",
    ci_high = "CI high"
  )
)

diagnostic_scenarios <- select_required_columns(
  diagnostic_scenarios_raw,
  c(
    "scenario",
    "paper_anchor",
    "outcome_family",
    "true_model_label",
    "fitted_model_label",
    "n_replications",
    "pseudo_interaction_detection_rate",
    "pseudo_interaction_detection_ci_low",
    "pseudo_interaction_detection_ci_high",
    "dharma_detection_rate_min",
    "dharma_detection_rate_max",
    "dharma_n_checks_used",
    "pregibon_detection_rate",
    "aic_favors_target",
    "quantitative_description",
    "variability_summary"
  )
)

diagnostic_scenarios <- rename_columns(
  diagnostic_scenarios,
  c(
    scenario = "Case",
    paper_anchor = "Paper anchor",
    outcome_family = "Outcome family",
    true_model_label = "Target model",
    fitted_model_label = "Misspecified model",
    n_replications = "Replications",
    pseudo_interaction_detection_rate = "Pseudo-interaction detection",
    pseudo_interaction_detection_ci_low = "Pseudo-interaction CI low",
    pseudo_interaction_detection_ci_high = "Pseudo-interaction CI high",
    dharma_detection_rate_min = "DHARMa min",
    dharma_detection_rate_max = "DHARMa max",
    dharma_n_checks_used = "DHARMa n checks",
    pregibon_detection_rate = "Pregibon check",
    aic_favors_target = "AIC favors target",
    quantitative_description = "Quantitative description",
    variability_summary = "Variability summary"
  )
)

diagnostic_scenario_display_table <- function(dat) {
  data.frame(
    Case = dat$Case,
    `Target model` = dat$`Target model`,
    `Misspecified model` = dat$`Misspecified model`,
    `Pseudo-interaction rate` = paste0(
      fmt_num(dat$`Pseudo-interaction detection`, 3), " ",
      fmt_ci_cols(dat$`Pseudo-interaction CI low`, dat$`Pseudo-interaction CI high`, 3)
    ),
    DHARMa = paste0(
      fmt_num(dat$`DHARMa min`, 3), "-",
      fmt_num(dat$`DHARMa max`, 3),
      " (", dat$`DHARMa n checks`, ")"
    ),
    Pregibon = ifelse(
      dat$Case == pregibon_na_scenario,
      "N.A.",
      fmt_num(dat$`Pregibon check`, 3)
    ),
    `AIC favors target` = fmt_num(dat$`AIC favors target`, 3),
    check.names = FALSE
  )
}

dg_lookup <- function(case, col) {
  if (!col %in% names(diagnostic_scenarios)) {
    stop(
      "Unknown diagnostic scenario column: ", col,
      ". Available columns: ",
      paste(names(diagnostic_scenarios), collapse = ", ")
    )
  }
  require_single(
    diagnostic_scenarios[diagnostic_scenarios$Case == case, col, drop = TRUE],
    paste0("diagnostic_scenarios[", case, " / ", col, "]")
  )
}
dg_pct <- function(case, col) round(100 * dg_lookup(case, col))
dg_n <- function(case) dg_lookup(case, "Replications")

# Replications attempted per diagnostic scenario, before any convergence loss.
n_rep_diagnostic <- max(table(diagnostic_replications$scenario))
dg_val <- function(case, col, digits = 1) fmt_num(dg_lookup(case, col), digits)
```

## Forced-Choice Accuracy and the Non-Zero Chance Floor

If a participant answers $Y_i$ of $k$ trials correctly, a binomial sampling model, $Y_i \sim \mathrm{Binomial}(k, p_i)$, is the natural starting point. The family requires the expected accuracy to lie between 0 and 1. When guessing is a plausible process for the task, the range that matters is narrower, since a participant who knows nothing still gets a share $c$ of the trials right, $c = .50$ in a two-alternative or true/false task and $c = 1/m$ with $m$ alternatives, so expected accuracy lies between $c$ and 1, even though observed proportions can fall below $c$ through sampling variation. Accuracy data of this kind are analyzed in several ways in practice, on the raw proportion scale with a Gaussian identity model, with a standard binomial logit or probit, and, less often, with a link that encodes the chance floor (@fig-motivating-example).

These specifications differ in which of the two ranges they impose. The Gaussian identity model imposes neither, since it ignores the trial process, assumes additivity in percentage-correct units, and can predict outside the 0-to-1 range. The other two keep the binomial family and differ only in the link. The standard logit or probit, $p_i = F(\eta_i)$ with $F$ the logistic or standard-normal CDF, holds expected accuracy between 0 and 1 and places the lower asymptote at 0, whereas the chance-corrected link maps the linear predictor onto the narrower interval from $c$ to 1,
$p_i = c + (1 - c)F(\eta_i)$, which for $c = .50$ becomes $p_i = .50 + .50\,F(\eta_i)$. 

An interaction term in the standard model therefore tests additivity over the full 0-to-1 range, whereas the same term in the chance-corrected model tests additivity on a performance-above-chance scale.

![A minimal $2 \times 2$ design crosses a condition (x-axis) with a group (color), shown on three scales. The four cell probabilities are identical in every panel; only the scale on which the interaction is evaluated changes. The data were generated with no condition-by-group interaction on the chance-corrected logit scale, which has a $.50$ chance floor, so the interaction coefficient is exactly zero on that scale. The same cells imply a positive interaction on the standard logit scale and a negative interaction on the raw probability scale. Light horizontal lines are equally spaced on each panel's own link scale, so equal steps on that scale map to unequal steps on the probability axis.](../figs/motivating-example.pdf){#fig-motivating-example}


The chance-corrected link is most defensible when the chance level is known from the task design, guessing is a plausible lower-bound process, and below-chance performance is not substantively meaningful for the target claim. In the simplest case, the chance level can be fixed by design, for example $c=0.50$ in a two-alternative task. In other cases, analysts may estimate the lower asymptote or report sensitivity analyses over plausible values of $c$. Lapses, response biases, item heterogeneity, changes across trials, or systematic below-chance performance may instead require richer psychometric solutions such as item-response models [e.g., @embretsonReise2000; @wichmann2001; @knoblauch2012]. For example, the 3-parameter logistic Item Response Theory model includes an item-specific guessing parameter rather than a single chance level fixed at the design level [@embretsonReise2000]. The chance-corrected link should thus be seen as a design-level approximation, not a substitute for item-level measurement models when the data and scientific question call for that additional structure.

### Simulation 1: Design and Results

```{r}
#| label: forced-choice-inline
#| include: false

# Design constants for Simulation 1 (see scripts/03a-simulation-forced-choice.R).
sim_N <- 250
sim_k <- 20
settings_age_low <- 6
settings_age_high <- 10
n_rep <- max(forced_choice_sim$`Successful fits`)

fc_lookup <- function(scn, mdl, col) {
  require_single(
    forced_choice_sim[
      forced_choice_sim$Scenario == scn & forced_choice_sim$Model == mdl,
      col, drop = TRUE
    ],
    paste0("forced_choice_sim[", scn, " / ", mdl, " / ", col, "]")
  )
}
fc_rejection_rate <- function(scn, mdl) round(100 * fc_lookup(scn, mdl, "Rejection rate"))
fc_interaction_sd <- function(
    scn,
    mdl = "Chance-corrected binomial logit"
) {
  fmt_num(
    fc_lookup(scn, mdl, "Monte Carlo SD interaction coefficient"),
    3
  )
}
matched_nominal_rate <- forced_choice_sim$`Rejection rate`[
  forced_choice_sim$Model == "Chance-corrected binomial logit"
]
matched_nominal_min <- round(100 * min(matched_nominal_rate))
matched_nominal_max <- round(100 * max(matched_nominal_rate))
```

In each of `r n_rep` replications we generated accuracy for $N = `r sim_N`$ participants with $k = `r sim_k`$ trials each, from a chance-corrected logit model with a .50 floor, a positive age effect, a group difference, and no age-by-group product term on the generating scale. Three scenarios differ only in overall performance, set by shifting the intercept: in the *lower-performance* scenario the two groups lie close to the .50 floor (expected accuracies roughly .53 to .80 across ages 6 to 10), in the *middle-performance* scenario they occupy the middle of the above-chance range (roughly .55 to .88), and in the *higher-performance* scenario they approach the ceiling (roughly .61 to .94). Each simulated dataset was analyzed with four models: a Gaussian linear model on the observed proportions, standard binomial-logit and binomial-probit GLMs on the correct-response counts, and the chance-corrected binomial-logit GLM that matches the generating scale. We recorded how often each product-term test rejected at $\alpha = .05$. The matching model supplies the nominal rejection rate, while the three alternative scales supply pseudo-interaction detection rates. @fig-forced-choice-simulation shows the simulation results; the per-scenario rates, fit-success counts, and coefficient variability are reported in Supplement A.

![Simulation 1: forced-choice accuracy with a chance floor. Data are generated with no age-by-group product term on the chance-corrected logit scale. Panel A shows the resulting accuracy curves across three performance levels; the grey region lies below the .50 chance floor. Panel B shows product-term rejection rates by fitted model and scenario. The matched chance-corrected model remains near nominal alpha, whereas the alternative scales produce pseudo-interactions. The dashed line marks alpha = .05.](../figs/forced-choice-simulation.pdf){#fig-forced-choice-simulation fig-pos="H"}

The chance-corrected model rejects between `r matched_nominal_min`% and `r matched_nominal_max`% of the time across the three scenarios. Pseudo-interaction detection is highest where performance lies near the floor: in the lower-performance scenario the standard logit rejects `r fc_rejection_rate("Lower performance", "Standard binomial logit")`% of the time, the standard probit `r fc_rejection_rate("Lower performance", "Standard binomial probit")`%, and the Gaussian identity `r fc_rejection_rate("Lower performance", "Gaussian identity")`%. Detection falls as performance moves away from the floor, reaching `r fc_rejection_rate("Higher performance", "Standard binomial logit")`% for the logit, `r fc_rejection_rate("Higher performance", "Standard binomial probit")`% for the probit, and `r fc_rejection_rate("Higher performance", "Gaussian identity")`% for the Gaussian identity in the higher-performance scenario. The same qualitative sensitivity to the chance floor also appears at chance levels 0.25 and 0.33 (Supplement B).

All four models returned a usable product-term test in every replication, including the lower-performance scenario. Complete fit success does not imply equal precision: under the matched chance-corrected model, the Monte Carlo standard deviation of the estimated age-by-group coefficient was `r fc_interaction_sd("Lower performance")` in the lower-performance scenario, compared with `r fc_interaction_sd("Middle performance")` in the middle-performance scenario and `r fc_interaction_sd("Higher performance")` in the higher-performance scenario. This greater sampling variability is consistent with less information about link-scale changes when expected accuracy lies closer to the chance floor. Because the four rejection rates are now computed on identical simulated datasets, the elevated rates under the alternative links cannot be attributed to selective convergence or Hessian failure.

## Within-Family Link Choice

The forced-choice case involved a link that was plainly wrong for the task, because it placed the lower asymptote at 0 when the design fixed it at chance. But the link-function problem can also arise in less obvious cases. Logit and probit are both standard for binary data, respect the same 0-to-1 bounds, and usually produce very similar fitted probabilities. Their choice is therefore often guided by convention, ease of interpretation, or implementation, with little expected impact on substantive conclusions [@yuhuang2019; @agresti2002]. This makes them a special case for illustrating the problem, as even two virtually interchangeable links define slightly different scales of additivity. Equal differences under one link are not exactly equal under the other, especially toward the tails, and this alone can alter the estimated product term.

The logit parameterization also gives coefficients a direct log-odds interpretation, whereas probit coefficients are expressed on a standard-normal latent scale, making it a choice of election when modeling psychological quantities such as ability, signal strength, and response tendency [@dunn2018]. 

### Simulation 2: Design and Results

```{r}
#| label: within-family-inline
#| include: false

# Design constants for Simulation 2 (see scripts/03b-simulation-within-family-links.R).
wf_N <- 700
wf_k <- 15
n_rep_wf <- max(within_family_sim$`Successful fits`)

wf_lookup <- function(gen, fit, col) {
  require_single(
    within_family_sim[
      within_family_sim$`Generating link` == gen & within_family_sim$`Fitted link` == fit,
      col, drop = TRUE
    ],
    paste0("within_family_sim[", gen, " / ", fit, " / ", col, "]")
  )
}
wf_rejection_rate <- function(gen, fit) round(100 * wf_lookup(gen, fit, "Rejection rate"))
wf_induced_product <- function(gen, fit) fmt_num(wf_lookup(gen, fit, "Induced product term"), 3)
```

In each of `r n_rep_wf` replications we generated repeated binary responses for $N = `r wf_N`$ participants in a two-group, two-condition design, with `r wf_k` trials per participant-condition cell and a subject random intercept (latent intraclass correlation 0.30). The quite large sample was chosen to make the relatively small systematic discrepancy between logit and probit links detectable, rather than to represent a typical psychological study. The condition-by-group interaction term was zero on the generating link scale. We crossed the generating link with the fitted link, using logit and probit in both roles, and fitted every model as a random-intercept GLMM, recording how often each returned a significant condition-by-group interaction term at $\alpha = .05$. @fig-within-family-links shows the results; the full crossed-link rates are tabulated in Supplement A. 

![Simulation 2: logit versus probit in repeated-trial random-intercept GLMMs. Data are generated with no condition-by-group product term on the generating link scale. Panel A shows the corresponding cell probabilities under logit and probit links. Panel B shows product-term rejection rates for each generating–fitted link combination. Matched links remain near nominal alpha, whereas wrong links produce pseudo-interactions. Error bars are Wilson 95% confidence intervals; the dashed line marks alpha = .05.](../figs/within-family-links.pdf){#fig-within-family-links fig-pos="H"}

When the fitted link matches the generating link, the nominal rejection rate stays near .05: `r wf_rejection_rate("logit", "logit")`% for logit fitted to logit-generated data and `r wf_rejection_rate("probit", "probit")`% for probit fitted to probit-generated data. Crossing the links raises the pseudo-interaction detection rate to `r wf_rejection_rate("probit", "logit")`% when a logit model is fitted to probit-generated data and `r wf_rejection_rate("logit", "probit")`% when a probit model is fitted to logit-generated data, roughly three times the nominal rate. The wrong-link detection rates are smaller than in the forced-choice case, and they appear among links that share the same family, the same 0-to-1 bounds, and nearly overlapping fitted curves in the middle of the probability range, the setting many applied analysts treat as interchangeable. Under the wrong link the same cell probabilities imply a small but deterministic nonzero product term on the fitted scale. Supplement B varies the intercept, which shifts where the four cell probabilities sit between 0 and 1, and shows that detectability changes non-monotonically with it, so the same link mismatch is nearly undetectable at some probability levels and much easier to detect at others. 

A result that is stable across both links is easier to defend; one that appears only under a single link should be reported as link-conditional. Supplement A, Section S5, reports how large that fixed product term is. Taking the cell probabilities implied by each generating model and re-expressing them on the other link scale gives a product term of `r wf_induced_product("logit", "probit")` for logit-generated data fitted with a probit and `r wf_induced_product("probit", "logit")` for probit-generated data fitted with a logit, against exactly zero whenever the links match. The rejection rates above are therefore the power to detect these effects.

## Sum Scores

Many psychological outcomes are computed as total sum scores across questionnaire items, task items, symptoms, errors, or ratings. In the two cases above the outcome was a directly observed response, and only the link was in question. A sum score is different, because the score itself is already the result of an analytic choice. What is observed are the individual item responses, which are binary or ordinal; the total is a summary of them, produced by a scoring rule that gives every item the same weight and every point of the scale the same width. In psychological research the construct of interest is often the underlying, generally continuous dimension, so the total can be seen as a remapped version of that continuum. The link question then arrives together with a prior question about which metric the claim is about.

The bounds are crucial here, because the construct may have no floor or ceiling while the observe total score has both. The mapping is best read as an exchange rate between the two scales, that is, how much of the latent dimension is needed to gain one raw-score point. That rate is not constant. Near the middle of the scale a small latent shift is enough to move the total by a point, while near either bound a wide stretch of the latent dimension is compressed into the last available point, because few items remain that the respondent can still change. To illustrate, with nine items scored 0-3, moving from 13 to 14 points requires less change on the latent dimension than moving from 26 to 27. A constant latent effect therefore produces smaller raw-score differences as the total approaches an end, and the same reasoning applies wherever the instrument is unevenly informative along the latent dimension, so compression is not confined to the endpoints. 

Bounded, discrete, and skewed distributions are common in psychological data [@micceri1989], and the consequences of analysing ordinal responses as if they were metric have been documented in detail [@liddellkruschke2018]. Fitting a Gaussian identity model to a total assumes instead that a predictor moves the score by the same number of points at every level of the other predictor. A group/condition sitting near a bound moves by fewer points even when the data-generating process on the latent dimension is the same in both groups/conditions. Behaviour-genetic research has met the same problem when testing whether genetic effects vary across environments, since those interactions are usually tested on summed questionnaire scores: analysed at the item level, the same data show that the total can conceal real interactions or present spurious ones through properties of the observed score [@molenaar2014]. This is the sum-score form of outcome-model misspecification and target-metric mismatch [@domingue2024].

The problem we raise sits next to a recent, wider debate about whether sum scores should be replaced by factor scores or latent-variable models. Some argue that sum scores are transparent, reliable enough for many purposes, comparable across studies, and tied to how instruments are scored in applied settings [@widaman2023]; the counterargument is that a total encodes strong and usually untested assumptions about how items represent the construct [@mcneish2020; @mcneish2023]. While we do not enter the debate, we add a narrower point that was left aside: a model of the manifest total defines the absence of an interaction in raw-score points, so an interaction found there is an interaction on the score, and it becomes a claim about the construct only under the additional assumption that the two are linearly related.

When item-level data are available and the claim concerns latent constructs, an explicit measurement model such as SEM, item-level ordinal regression, or IRT-inspired specifications could be preferred, since it addresses the mapping directly, though at the price of moving the assumptions into the measurement model [@burknervuorre2019]. Often, however, researchers deliberately work with the total score for the practical reasons listed above. An option is to set a family that respects the bounds and the discreteness, for example modelling the total as a bounded count with a beta-binomial likelihood [@wilcox1981]. However, we suggest a deliberately unusual option that moves only the link: keep the observed total score under a Gaussian likelihood, but rescale the score to the proportion of the maximum and use a probit link function, as in Simulation 3 below. This is an *ad hoc* device, not a standard practice, but it illustrates whether the interaction conclusion changes if additivity is placed on a nonlinear scale that compresses the expected values toward the score bounds. It remains a model of the total, it will not recover the construct, and the Gaussian likelihood is still misspecified in its variance, since a bounded score has a variance that depends on its mean: what the probit link corrects is only the scale on which additivity is defined, not the distributional form. Read against the sum-score debate, it is a precaution for researchers who have decided to analyse totals anyway, but acknowledge that the interaction still needs a scale on which additivity is plausible.

### Simulation 3: Design and Results

```{r}
#| label: sum-score-inline
#| include: false

# Design constants for Simulation 3 (see scripts/03c-simulation-sum-scores.R).
ss_N   <- 600
ss_J   <- 9
ss_cat <- 3                 # each item scored 0 to 3
ss_max <- ss_J * ss_cat     # 0-27 total
n_rep_ss <- max(sum_score_sim$`Successful fits`)

ss_lookup <- function(scn, mdl, col) {
  require_single(
    sum_score_sim[
      sum_score_sim$Scenario == scn & sum_score_sim$Model == mdl,
      col, drop = TRUE
    ],
    paste0("sum_score_sim[", scn, " / ", mdl, " / ", col, "]")
  )
}
ss_rejection_rate <- function(scn, mdl) round(100 * ss_lookup(scn, mdl, "Rejection rate"))

# Ranges quoted in the text are the observed min/max across scenarios, so they
# stay ordered if a re-run changes which scenario is the extreme one.
ss_rate_range <- function(mdl) {
  rates <- sum_score_sim$`Rejection rate`[sum_score_sim$Model == mdl]
  if (length(rates) == 0) stop("No rejection rates found for model ", mdl, ".")
  round(100 * range(rates))
}
ss_rate_min <- function(mdl) ss_rate_range(mdl)[1]
ss_rate_max <- function(mdl) ss_rate_range(mdl)[2]
```

The bounded-score probit model was compared with two natural alternatives: the observed total and the latent scale it approximates. In each of `r n_rep_ss` replications, we generated data for $N = `r ss_N`$ participants from a continuous latent variable with a continuous predictor $x$, a group effect, and no x-by-group interaction. The latent values generated `r ss_J` ordinal items scored 0 to `r ss_cat`, which were summed to a total ranging from 0 to `r ss_max`. Item thresholds were shifted across three scenarios so that expected totals fell in the middle, lower, or upper part of the score range.

We analyzed each total in three ways: a Gaussian identity model of the observed sum score, a Gaussian-probit bounded-score model, and the true latent generating scale. The bounded-score model applies a probit link to a continuity-corrected score proportion and rescales predictions to the original sum-score metric. We used probit because its standard-normal linear predictor provides a simple correspondence with the latent scale. This choice is still substantive, as shown in the previous section. The latent model serves as an idealized benchmark available only in simulation. We recorded the proportion of significant x-by-group interactions at $\alpha = .05$. Results are shown in @fig-sum-score-simulation, with per-scenario rates in Supplement A, Section S4.

![Simulation 3: sum scores as bounded and discrete outcomes. Data are generated with no x-by-group product term on the latent scale. Panel A shows how the same additive latent structure maps onto sum scores in the middle, lower, and upper parts of the score range. Panel B shows product-term rejection rates for the latent benchmark and two observed-score models. The latent model remains near nominal alpha, whereas the observed-score models can produce pseudo-interactions. The dashed line marks alpha = .05. The latent benchmark is available only in simulation.](../figs/sum-score-simulation.pdf){#fig-sum-score-simulation fig-pos="H"}

The latent benchmark remained close to the nominal .05 rejection rate in every scenario, from `r ss_rate_min("Latent generating scale")`% to `r ss_rate_max("Latent generating scale")`%. The identity model behaved similarly only in the middle of the score range, where the sum score was approximately linear in the latent variable (`r ss_rejection_rate("Middle range", "Observed sum-score identity")`%). Rejection increased as expected totals moved toward a bound, reaching `r ss_rejection_rate("Upper range", "Observed sum-score identity")`% in the upper range and `r ss_rejection_rate("Lower range", "Observed sum-score identity")`% in the lower range. Compression of the latent scale makes the same latent group difference appear to change across levels of the predictor. The bounded-score probit model remained close to nominal alpha across scenarios, from `r ss_rate_min("Gaussian-probit bounded score")`% to `r ss_rate_max("Gaussian-probit bounded score")`%. It respects the score bounds, although it is still a model of the manifest total.

This pattern mainly depends on where observations fall within the score range. Across the broader scenarios in Supplement B, the position of expected scores within the range affected identity-model rejection much more than sample size, item count, or latent dispersion. Raw-score additivity can approximate latent-scale additivity when scores remain away from the bounds. Compression near a bound can instead create an apparent interaction from a constant latent effect. The bounded-score probit model can therefore serve as a useful sensitivity analysis. For claims about interactions between latent constructs, an explicit measurement model provides a more direct approach, for example SEM, item-level ordinal models, or IRT-inspired models. These approaches still depend on assumptions about thresholds, links, loadings, and latent-response scales, but they make those assumptions part of the interaction model.

# Can Standard Diagnostics Catch a Wrong Link?

The simulations above assume the analyst does not question the fitted model. In practice, a researcher who finds an interaction may run model checks before reporting it. We examine the three checks an applied psychologist is most likely to run, each representing a different strategy. A Pregibon-style added-term test probes the assumed link, residual diagnostics screen for misfit of any kind, and AIC ranks a set of candidate models. Estimating the shape of the link from the data through a parametric family of links is a more direct strategy, and it is rarely used in psychology [@aranda-ordaz1981; @stukel1988; @czadosantner1992]. None of them identifies the scale on which the no-interaction claim is defined. At best they point to the link that fits the data best among the alternatives supplied. 

A link test of the kind proposed by Pregibon targets link adequacy directly, in the added-variable form used here [@pregibon1980]. The model is fitted, the fitted linear predictor $\hat{\eta}$ is extracted, and the model is refitted with $\hat{\eta}^2$ added as an extra predictor. If the link is adequate the squared term carries no additional information and should be non-significant; a significant term signals curvature that the current link leaves unexplained. The signal is not specific to the link, as an omitted nonlinear term, an omitted covariate, or a missing interaction produce it as well. A non-significant result means only that this diagnostic did not detect a departure in this sample.

Residual diagnostics start from the position of each observed value within the distribution the model predicts for that observation. That position is a number between 0 and 1, and under an adequate model those positions are uniform whatever the family, which gives residual plots a common reference distribution and makes misfit visible as departure from uniformity, as over- or underdispersion, or as residual structure over fitted values and predictors. Dunn and Smyth pass the position through the normal quantile function, so their quantile residuals are standard normal under an adequate model [@dunnsmyth1996], while DHARMa obtains the same positions by simulating datasets from the fitted model and leaves them on the 0-to-1 scale [@hartig2024]. We use the DHARMa implementation in R. As with the link test, several causes can produce departure from uniformity, including link misspecification, wrong family, missing nonlinearity, missing interaction, and target-metric mismatch. 

Information criteria compare a supplied set of models. We use AIC throughout, and because our candidate models keeps the response, family, and structural formula fixed and varies only the link, the ranking depends entirely on the log-likelihood.  Importantly, the result is restricted to the candidate links that were fitted, and the interaction term is already present in each model and can absorb part of the curvature produced by the wrong link. A misspecified model may therefore fit almost as well as the target model while still placing additivity on the wrong scale. 

## Diagnostic Simulation

The five scenarios reuse examples already introduced in the paper (@tbl-diagnostic-scenarios). Each keeps the outcome family and the interaction formula fixed and changes only the link, so the fitted model is wrong about the scale on which effects add and about nothing else. *Count errors* reproduces the Poisson-log example of @fig-simple-count-example and fits it with an identity link. *Chance-floor accuracy* takes the lower-performance scenario of Simulation 1, generated on the chance-corrected logit scale and fitted with a standard binomial logit. *Binary repeated trials* reproduces the probit-generated, logit-fitted case of Simulation 2. *Binary continuous predictor* repeats that case with the two-level condition replaced by a continuous predictor, which allows the residual checks that need variation along a predictor. *Gamma mean response time* generates response times from a Gamma model with a log link and fits the canonical inverse link. The generating parameters for all five are reported in Supplement A.

In each replication, diagnostics were applied to the misspecified interaction model, because this is the model an analyst would usually inspect after obtaining a significant interaction. All rates are computed over the replications in which that model converged, which is every one of the `r n_rep_diagnostic` replications except in the count scenario, where the Poisson-identity model returned a usable fit in `r dg_n("Count errors")` of them. The question is therefore whether the same model that produces the pseudo-interaction also shows enough evidence of misspecification to warn the analyst. 

Each replication was screened with four DHARMa checks, covering residual uniformity, dispersion, residual quantile patterns over fitted values, and residual structure over the focal predictor [@dunnsmyth1996; @hartig2024]. The last of these takes two forms, a quantile check when the focal predictor is continuous and a comparison across design cells in the categorical $2 \times 2$ case. Alongside them we ran the Pregibon-style added-term link check [@pregibon1980]. Every procedure was applied a second time to the correctly specified interaction model fitted to the same data, which calibrates the flagging rates against a model that is right.

The AIC comparison in each scenario is between the two models already described, the target and the misspecified one. We report the proportion of replications in which AIC favored the target model.

```{r}
#| label: tbl-diagnostic-scenarios
#| tbl-cap: "Diagnostic simulation across five link-misspecification scenarios. Pseudo-interaction reports how often the misspecified model detected an interaction that was absent on the generating link scale. DHARMa gives the range across the four prespecified residual checks. Pregibon reports the added-term link check. AIC favors target reports the proportion of replications in which the lower AIC favored the target interaction model over the misspecified interaction model. CI = Wilson 95% confidence interval. The Pregibon check is not identifiable in the saturated $2 \\times 2$ design, where the squared linear predictor is exactly collinear with the existing fixed effects."
#| echo: false

knitr::kable(
  diagnostic_scenario_display_table(diagnostic_scenarios),
  align = c("l", "l", "l", "r", "r", "r", "r")
)
```

@tbl-diagnostic-scenarios shows that the three checks behave very differently. The Pregibon test flagged the count case in `r dg_pct("Count errors", "Pregibon check")`% of replications, but reached only `r dg_pct("Gamma mean response time", "Pregibon check")`% in the Gamma case and `r dg_pct("Chance-floor accuracy", "Pregibon check")`% at the chance floor, fell to `r dg_pct("Binary continuous predictor", "Pregibon check")`% for the binary case with a continuous predictor, and could not be computed at all in the saturated $2 \times 2$ design. The DHARMa checks were weaker, reaching `r dg_pct("Count errors", "DHARMa max")`% in the count case and at most `r dg_pct("Chance-floor accuracy", "DHARMa max")`% in the other four, so residual screening responded only where the wrong link visibly distorted the fitted values. AIC favored the target link in every scenario, from `r dg_pct("Gamma mean response time", "AIC favors target")`% in the Gamma case to `r dg_pct("Count errors", "AIC favors target")`% in the count case, and was the only check that responded throughout.

The two logit-probit cases were the hardest to diagnose. They produced the fewest pseudo-interactions, `r dg_pct("Binary repeated trials", "Pseudo-interaction detection")`% in the saturated design and `r dg_pct("Binary continuous predictor", "Pseudo-interaction detection")`% with a continuous predictor, and almost nothing detected them, leaving AIC at `r dg_pct("Binary repeated trials", "AIC favors target")`% and `r dg_pct("Binary continuous predictor", "AIC favors target")`% as the only informative signal.

The chance-floor scenario is worth a closer look, because it combines a high pseudo-interaction rate with residual checks that are almost never significant, which leaves AIC as the only diagnostic that detects the misspecification. There AIC favored the chance-corrected model in `r dg_pct("Chance-floor accuracy", "AIC favors target")`% of replications, but the margin is often narrow, and Supplement A reports the full distribution of the AIC difference for every scenario.

Overall, the diagnostics were informative in some settings and weak in others, even when the pseudo-interaction significance rate is pretty high. Under the correct link, DHARMa flagged a problem in at most `r fmt_num(dharma_calibration_max, 1)`% of replications. Pregibon flagged at most `r fmt_num(pregibon_calibration_single_level_max, 1)`% in the count, chance-floor, and Gamma scenarios, and `r fmt_num(pregibon_calibration_glmm_continuous, 1)`% in the continuous-predictor GLMM.

Full results are reported in Supplement A, and Supplement B shows how pseudo-interaction rates, Pregibon detection, and AIC preference change across sample size and the intercept.

# From Link Choice to Sensitivity Analysis

The practical implication of this framework is that an interaction analysis should make three choices explicit: the link that defines additivity, the metric the claim refers to, and the sensitivity checks used to assess that choice (@tbl-link-aware-workflow). The observed metric may sometimes be the relevant target. For example, a two-point difference on a clinical or educational score may be the quantity that matters in practice, and additivity in raw-score units may then be reasonable. The key point is that this should be a substantive choice, not the by-product of using a default model. When possible, the link should be justified from theory, task structure, or measurement [@baranger2023; @castillo2026].

When link choice can  affect the interaction conclusion, we suggest that the main robustness check is to performa a sensitivity analysis across plausible links. Researchers should therefore compare the same target response-scale contrast across alternatives and report whether the result is stable or link-conditional. If it changes, that dependence is itself part of the result [@mccabe2022; @mize2019; @rohrer2026; @mustillo2018].

```{r}
#| label: tbl-link-aware-workflow
#| tbl-cap: "Practical starting points for link-aware interaction testing across common psychological outcomes."
#| tbl-colwidths: [20, 42, 38]
#| echo: false

workflow <- data.frame(
  `Outcome type` = c(
    "Forced-choice accuracy",
    "Binary / proportion outcome",
    "Likert / ordinal item",
    "Sum score / composite",
    "Counts / error counts",
    "Response time / skewed positive outcome"
  ),
  `Plausible starting point` = c(
    "Binomial logit/probit; chance-corrected link when the task floor is part of the estimand.",
    "Binomial logit/probit; consider cloglog for asymmetric processes.",
    "Ordinal logit/probit; metric model only with a clear observed-scale target.",
    "Gaussian identity when many levels stay far from bounds; otherwise consider item-level, ordinal, bounded-count, or bounded-score models.",
    "Poisson or negative binomial, usually with log link; identity only with an observed-count target.",
    "Lognormal, Gamma log/inverse, Inverse Gaussian, or transformed Gaussian, matched to absolute, proportional, reciprocal, or transformed change."
  ),
  `Minimal sensitivity check` = c(
    "Compare standard logit/probit and chance-corrected links; report convergence and response-scale contrasts.",
    "Compare logit/probit; inspect calibration and contrasts near bounds or rare-event regions.",
    "Compare ordinal and metric summaries; report category probabilities or response-scale contrasts.",
    "Inspect bounds, skew, and score concentration; compare with item-level or bounded-score sensitivity models when feasible.",
    "Check overdispersion and zero inflation; compare count contrasts across plausible count models.",
    "Compare raw-, log-, and response-scale summaries; inspect tail fit and predicted values."
  ),
  check.names = FALSE
)

knitr::kable(workflow, align = c("l", "l", "l"))
```

# Discussion

The same data can support different interaction conclusions under different plausible specifications, which is the familiar scale dependence of interactions [@cox1984; @loftus1978; @wagenmakers2012]. In a GLM that dependence is carried by two choices, the outcome family and the link function. Our review found interaction tests to be common in all five journals sampled, the link function usually left implicit, and identity-link analyses applied to many outcomes for which observed-scale additivity is not self-evident.

The simulations then showed what that can cost. With two main effects present, a link that misplaces additivity turned a purely additive generating process into a significant product term even when the outcome family was correct. In forced-choice accuracy the binomial family is correct and the error lies entirely in a canonical link that places the lower asymptote at 0 rather than at the chance level. The logit-probit comparison showed that the problem survives when the response bounds are correct as well, between two links that look alike over most of the range. In the sum-score case the choice extends beyond the link to the outcome itself. A total is a scoring rule applied to item responses, so an interaction estimated on it refers to raw-score points, and adopting a family that respects the bounds and the discreteness of the total still leaves open whether additivity in points stands for additivity in the construct.

This difficulty is of a different kind from the one documented in work on the precision of interaction estimates [@baranger2023; @castillo2026]. There the obstacle is noise, and more data helps. Here the product term induced by the wrong link is a fixed property of that scale, so a larger sample estimates it more precisely and reports it as significant more often.

The diagnostic simulation showed that ordinary post-fit checks catch this unevenly. Residual checks, Pregibon-style tests, and AIC each detected the misspecification in some scenarios and not in others, and their sensitivity depended on which candidate models were supplied. The chance-floor scenario illustrates this most clearly, since a wrong link produced pseudo-interactions in most replications while the residual checks were almost never significant, and the only check that detected the misspecification was the explicit comparison between the two candidate links, which preferred the correct link by a narrow margin. These diagnostics therefore answer a question about fit within a supplied candidate set, which is a different question from whether an interaction claim holds on the scale the researcher means [@dunnsmyth1996; @hartig2024; @pregibon1980].

Three limitations bound the scope of these conclusions. The review was descriptive and covered five reference journals across major areas of psychology in a single year, so its estimates describe reporting practice in high-standard outlets rather than the field as a whole, and a broader mapping across journals, years, and subfields would be needed to support reporting guidance. The review also coded observable reporting features and did not judge, article by article, whether a given interaction was scale-dependent enough to be removable. The simulations, finally, are demonstrations of a mechanism. They target a single regime, in which two established main effects push expected values into the region where link curvature is strongest, and they say nothing about how often that regime arises or about the further complications of real data.

To conclude, interaction claims rest on a scale that is rarely stated. Naming it separates three questions that a software default otherwise settles at once, which family suits the response, which link defines additivity, and which metric the claim is about. The robustness question then takes a definite form, whether the conclusion survives the links that remain defensible. A result that holds across them is robust to that choice. One that does not is still a result, and it should be reported as conditional on the scale on which it was estimated.

# References

::: {#refs}
:::
