---
title: "Supplementary Materials for The Link Function Problem in Psychological Interaction Testing"
bibliography: references.bib
format:
  html: default
  pdf: default
---

# Review Methods

This supplement reports details of the preregistered descriptive review, mathematical background, simulation data-generating processes, and reproducibility information. The main text reports only compact summaries.

## Protocol and Scope

Describe the preregistered protocol, selected journals, search source, inclusion and exclusion criteria, and screening workflow. Include the OSF or repository link once available.

## Coding Variables

Summarize the variables from the review data dictionary: interaction testing, non-identity link use, explicit link reporting, potentially problematic identity-link analyses, significant interactions, and response-variable categories.

## Additional Review Tables

```{r review-supplement-tables}
#| echo: false
#| results: asis

journal_table_path <- "../tables/tableS1-review-by-journal.csv"
if (file.exists(journal_table_path)) {
  journal_table <- read.csv(journal_table_path)
  knitr::kable(journal_table, caption = "Review descriptives by journal.", digits = 1)
} else {
  cat("Run scripts/01-review-descriptives.R to generate the journal-level review table.")
}
```

# Mathematical Details

## GLM Notation

A GLM links the expected outcome $\mu_i = E(Y_i)$ to a linear predictor $\eta_i$ through

$$
g(\mu_i) = \eta_i.
$$

With two predictors and a product term,

$$
\eta_i = \beta_0 + \beta_1 X_i + \beta_2 Z_i + \beta_3 X_iZ_i.
$$

The coefficient $\beta_3$ is an interaction on the link scale. Response-scale interactions require applying the inverse link and computing the relevant contrast.

## Response-Scale Difference-in-Differences

For binary or categorical predictors, one response-scale interaction estimand is

$$
\{E(Y \mid X = 1, Z = 1) - E(Y \mid X = 0, Z = 1)\}
-
\{E(Y \mid X = 1, Z = 0) - E(Y \mid X = 0, Z = 0)\}.
$$

In nonlinear GLMs this quantity can differ from the product-term coefficient in sign, size, or statistical conclusion.

# Chance-Corrected Link Details

The simple fixed-chance link used in the paper is

$$
p_i = c + (1 - c)F(\eta_i),
$$

where $c$ is the chance level and $F$ is a cumulative distribution function such as the logistic or normal CDF. For 2-AFC tasks, $c = .50$.

A more general version with a lapse parameter is

$$
p_i = c + (1 - c - \lambda)F(\eta_i),
$$

where $\lambda$ controls the upper asymptote. The main paper uses the simpler fixed-chance version for transparency.

# Simulation Data-Generating Processes

## Simulation 1: Forced-Choice Accuracy

Report the settings used in `scripts/03a-simulation-forced-choice.R`: number of replications, sample sizes, trials per subject, predictor distributions, generating coefficients, fitted models, and false-positive criterion.

## Simulation 2: Sum Scores

Report the latent-variable model, item thresholds, number of items, response categories, fitted models, and false-positive criterion used in `scripts/03b-simulation-sum-scores.R`.

## Simulation 3: Within-Family Links

Report the binary outcome DGP, generating links, fitted links, and scenario definitions used in `scripts/03c-simulation-within-family-links.R`.

## Diagnostic Worked Example

Report the diagnostic scenarios, the Pregibon-style added-term check, the false-positive interaction criterion, and the diagnostic detection criterion used in `scripts/04-diagnostic-worked-example.R`.

# Extended Simulation Results

```{r extended-simulation-results}
#| echo: false
#| results: asis

files <- c(
  "../tables/table3-forced-choice-simulation.csv",
  "../tables/table4-sum-score-simulation.csv",
  "../tables/table5-within-family-links.csv",
  "../tables/table6-diagnostic-worked-example.csv"
)

for (f in files) {
  if (file.exists(f)) {
    cat("\n\n")
    knitr::kable(read.csv(f), caption = basename(f), digits = 3)
  }
}
```

# Reproducibility

All simulations set seeds in the corresponding scripts. Heavy computation is kept outside the manuscript file. The main paper reads saved tables and figures.

```{r session-info}
#| echo: false
sessionInfo()
```
